# Linked Micromap Plots via the **micromap** R Package {#Ch2}


\chapterauthor{J{\"u}rgen Symanzik, Marcus W. Beck, Michael McManus}


The **micromap**\index{R Packages!micromap} R package [@PaOl2015],
accessible at https://cran.r-project.org/web/packages/micromap/index.html,
will be introduced in this chapter. The reader will learn how to create a 
basic linked micromap plot\index{Linked micromap plot} via this R package. 
Details will be provided how to optimize and fine-tune such a basic plot 
into a publication-worthy final linked micromap plot.


## Introduction {#Ch2-Introduction}


As a reminder, see Chapter \@ref(Ch1) for general style requirements
for our `Micromap Plots in R` book. In particular, please do the following:

- Introduce meaningful labels for the sections, figures, and tables in your chapter.

- Create index entries for all R packages (such as the **micromap**\index{R Packages!micromap} R package)
and for all datasets (such as the _USstates_\index{Datasets!USstates} and _edPov_\index{Datasets!edPov} datasets)
that are used in your chapter.

- Include references for R packages and publications related to your chapter,
such as for the **micromap**\index{R Packages!micromap} [@PaOl2015] and 
**micromapST**\index{R Packages!micromapST} [@CP2015CRAN] R packages
and some micromap articles, book chapters, and books [@Carr2001;@SC2008;@CP2010].

- Also create index entries for main topics such as
linked micromap plots,\index{Linked micromap plot}
conditioned choropleth maps,\index{Conditioned choropleth map}
perceptual group,\index{Perceptual group}
color blindness,\index{Color blindness},
and quantile-quantile plot.\index{Quantile-quantile plot}


## Main {#Ch2-Main}


Here goes the main content of your chapter. Introduce additional sections as needed.

For convenience, Figure \@ref(fig:Ch2-micromap1) shows one linked micromap plot\index{Linked micromap plot}
(which is the same as in Figure \@ref(fig:Ch1-micromap1)), but now formatted in a slightly more meaningful way.


```{r Ch2-micromap1, fig.cap = 'Here is a first micromap example for this chapter. Note that the figure is formatted in a slightly more meaningful way this time.', fig.width = 7, fig.height = 9}
library(micromap)

# initial example

data(USstates)
statePolys <- create_map_table(USstates, "ST")
data(edPov)

# basic figure 1
lmplot(
  stat.data = edPov,
  map.data = statePolys,
  panel.types = c("labels", "dot", "dot", "map"),
  panel.data = list("state", "pov", "ed", NA),
  ord.by = "pov",   
  grouping = 5, 
  median.row = TRUE,
  plot.width = 2, 
  plot.height = 6,
  map.link = c("StateAb", "ID")
)
```


## Chapter Outline {#Ch2-Outline}


Note: Chapter \@ref(Ch1) will cover elements of a linked micromaps, 4 panels of legend, label, statistic, micromaps 
and rows of perceptual groups? Also discussion of spatial patterns and interpretation of positive, negative
(or no) association in statistical panels of dot plots.


-	Overview of Four Steps (**Marcus**: Maybe a good figure could be a conceptual diagram showing the whole process?)
    - Identifying and Geoprocessing of Spatial Boundary Data
    - Linking Spatial Boundary Data and Statistical Data
    - Creating a Draft Linked Micromap Plot
    - Refining the Linked Micromap Plot
    
-	Spatial Polygons in R and micromap Package Updates (**Juergen**: How much of this do we need here in Chapter \@ref(Ch2)? Would it make more sense to mention this in Chapter \@ref(Ch4)?)
    - Spatial objects and simple feature objects
    - Updates
    
-	Identifying and Geoprocessing of Spatial Boundary Data (**Marcus**: I think this is a good topic to discuss, but may be tricky to balance specificity with brevity.  I'm no expert on it, but I know there are a few different methods one could use.  Maybe just hit on the why and how, and briefly present tradeoffs of the different methods?  I think for MM, the goal is just to increase render time w/o sacrificing spatial specificity, so maybe the exact method is not important? -- **Juergen**: In fact, let's keep it simple here and work with ready-to-use shapefiles that are already available in some R packages; refer to Chapter \@ref(Ch4) how to bring in external shapefiles and modify them for use in micromaps)
    - Keep this brief as Chapter \@ref(Ch4) might cover this in detail?
    - Example of generalizing polygons Figure  (Use same figures as Chapter \@ref(Ch4)?)

-	Linking Spatial Boundary Data and Statistical Data
    - Explicit link to spatial object
    - Under the hood link to simple features object

-	Creating a Draft Linked Micromap Plot
    - Use workshop NARS pH sample & population estimates Figure (**Juergen**: Use simpler edPov data in first example)

-	Refining the Linked Micromap Plot
    - Use workshop NARS pH sample & population estimates Figure (**Juergen**: Use simpler edPov data in first example)

- **Juergen**: Introducing a second basic example for a different geographic region, e.g., state level or environmental, and use of a different built-in plot type (e.g., boxplot). Similar four steps as before, but more condensed.

-	Challenges to Visualizing Data with Linked Micromap Plots (**Marcus**: Not sure this is the right place for it, but maybe it's worth providing a brief contrast with micromapST?  Maybe include this in the overview at the top?  I think it would be useful for the reader to understand succinctly how the two differ and why there are two packages, though I'm sure these themes will be present throughout the book.  Perhaps the latter point is not that interesting, i.e., it's common to have "competing" software in an open source environment? -- **Juergen**: Yes, we need some discussions at the end of Chapters \@ref(Ch2) and \@ref(Ch3) that outline advantages and limitations of each of the two main micromap packages. Alternatively, there could be some table in Chapter \@ref(Ch1) that lists features (and limitations) of the two packages when they are first introduced to the reader. Overall, this last point may be a subsection in the Introduction, or become a final chapter at the end of the book which research is necessary in the future.)
    - Problem of many polygons (**Juergen**: There are other problems, e.g., many islands as in the Philippines, narrow shapes as for Chile, subregions of considerably different sizes as for Russia)
    - Automating linked micromap plot production for reports and interactive visualizations (Check with Emma Jones at VA DEQ)
    - (**Marcus**: Related, a disadvantage of both packages is their lack of tighter integration with ggplot.  Not sure we want to go there though. -- **Juergen**: I wouldn't go there as these are the packages that currently exist and have to be used.)


## Steps to Create a Linked Micromap Plot with the **micromap**\index{R Packages!micromap} R Package {#Ch2-Steps}


Four main steps, shown in Figure \@ref(fig:Ch2-flowchart), are needed to create a 
linked micromap plot\index{Linked micromap plot} 
with the **micromap**\index{R Packages!micromap} R package:

1. **Identifying and Geoprocessing of Spatial Boundary Data**: In addition to a data frame that contains the statistical data,
the user must identify a data structure that contains the spatial boundary data for the region and subregions that have
to be colored in the graph. This spatial boundary data typically comes in the form of a
SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame}. In this chapter, we will work with ready-to-use
boundary files. In Chapter \@ref(Ch4), we will discuss how to make use of boundary files that are
provided as external shapefiles\index{Shapefiles}. We will also see in Chapter \@ref(Ch4) how
to simplify complex boundaries, enlarge small subareas in the maps, and move subareas that are far away closer
to the main area. Eventually, the SpatialPolygonsDataFrame has to be transformed into a regular data frame
via the `micromap::create_map_table()` function.

2. **Linking Spatial Boundary Data and Statistical Data**: Next, we have to identify variables from the statistical data frame
and the newly created data frame with the boundary information that that allow us to link statistical data and boundary data
for each subregion.

3. **Creating a Draft Linked Micromap Plot**: While not necessary, it is always a good idea to create a minimal
linked micromap plot\index{Linked micromap plot} to ensure that the statistical data and boundary data
are matching and a correct linked micromap plot\index{Linked micromap plot} is created. Skipping this
step and trying to create a complex linked micromap plot\index{Linked micromap plot} may make it hard
to find possible errors in the R code.

4. **Refining the Linked Micromap Plot**: Once a draft linked micromap plot\index{Linked micromap plot} has been
created, it usually needs fine-tuning of its appearance and plot aesthetics, e.g., modification of colors,
layouts, and groups of areas, addition of labels and legends, and possibly the additional of additional
statistical variables or change to a different graph type.


```{r Ch2-flowchart, fig.cap = 'Workflow to create a linked micromap plot with the **micromap**\\index{R Packages!micromap} R package (Diagram created with the **DiagrammeR**\\index{R Packages!DiagrammeR} R package).', fig.width = 5, fig.height = 5, echo = FALSE}
library(DiagrammeR)

grViz(diagram = "digraph flowchart {
  node [fontname = arial, shape = oval, color = grey, style = filled]
  tab1 [label = '@@1']
  tab2 [label = '@@2']
  tab3 [label = '@@3']
  tab4 [label = '@@4']

  tab1 -> tab2 -> tab3 -> tab4;
}

  [1]: '1. Identifying and Geoprocessing of Spatial Boundary Data'
  [2]: '2. Linking Spatial Boundary Data and Statistical Data'
  [3]: '3. Creating a Draft Linked Micromap Plot'
  [4]: '4. Refining the Linked Micromap Plot'
")
```



### Identifying and Geoprocessing of Spatial Boundary Data {#Ch2-Identifying}

In this first example, we work with the
_USstates_\index{Datasets!USstates} and _edPov_\index{Datasets!edPov} datasets
from the **micromap**\index{R Packages!micromap} R package.
_USstates_\index{Datasets!USstates} is a SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame} for 
the 50 states (and Washington, D.C.) of the United States (U.S.) 
that was created for use in linked micromap plots\index{Linked micromap plot}. Notably, the boundaries
of many of the states have been simplified, Alaska and Hawaii have been moved closer to the
contiguous 48 states (and also have been resized), and Washington, D.C., has been pulled out of the
main map and has been placed to the east (and also has been enlarged).
_edPov_\index{Datasets!edPov} is a data frame that contains
education and poverty level data for the 50 states (and Washington, D.C.).



### Linking Spatial Boundary Data and Statistical Data {#Ch2-Linking}


### Creating a Draft Linked Micromap Plot {#Ch2-Creating}


### Refining the Linked Micromap Plot {#Ch2-Refining}




## Code Chunks {#Ch2-CodeChunks}


Example code chunks from Mike's workshop materials. Reduce some of the figures.


```{r Ch2-micromap2, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
#I. Overview of Four Steps to Make a Linked Micromap

### I. A. Geoprocessing of Spatial Data
library(micromap)
sessionInfo()
getwd()

library(labeling)

data("USstates")
#Here is the spatial data
plot(USstates)
head(USstates@data)
```



```{r Ch2-micromap3, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### I. B. Structure:  Create a Map Table to Link to the Statistical Data
statePolys <- create_map_table(USstates, 'ST')
#The spatial data is now in a simpler form that the micromap package can use
head(statePolys)

data("edPov")
#Here is the statistical data.
head(edPov)
#The spatial and statistical data frames must have a common linking variable.
hist(1:100)
```



```{r Ch2-micromap4, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### I. C. Draft Micromap Plot
mmplot(stat.data=edPov,
       map.data=statePolys,
       map.link=c('StateAb', 'ID'),
       panel.types=c('dot_legend', 'labels', 'dot', 'dot', 'map'),
       panel.data=list(NA, 'state', 'pov', 'ed', NA),
       ord.by='pov',   
       grouping=5, median.row=TRUE)
```



```{r Ch2-micromap5, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### I. D. Basic Modification Exercises of Draft Micromap Plot

### I. D. a) Reorder the previous plot in decreasing order of poverty
mmplot(stat.data=edPov,
       map.data=statePolys,
       map.link=c('StateAb', 'ID'),
       panel.types=c('dot_legend', 'labels', 'dot', 'dot', 'map'),
       panel.data=list(NA, 'state', 'pov', 'ed', NA),
       ord.by='pov',   
       rev.ord=TRUE,
       grouping=5, median.row=TRUE)
```



```{r Ch2-micromap6, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### I. D. b) Change the order of the education and poverty panels and sort by decreasing education
mmplot(stat.data=edPov,
       map.data=statePolys,
       map.link=c('StateAb', 'ID'),
       panel.types=c('dot_legend', 'labels', 'dot', 'dot', 'map'),
       panel.data=list(NA, 'state', 'ed', 'pov', NA),
       ord.by='ed',   
       rev.ord=TRUE,
       grouping=5, median.row=TRUE)
```



```{r Ch2-micromap7, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### I. D. c) Place the maps on the left side
mmplot(stat.data=edPov,
       map.data=statePolys,
       map.link=c('StateAb', 'ID'),
       panel.types=c('map', 'dot_legend', 'labels', 'dot', 'dot'),
       panel.data=list(NA, NA, 'state', 'ed', 'pov'),
       ord.by='ed',   
       rev.ord=TRUE,
       grouping=5, median.row=TRUE)
```



```{r Ch2-micromap8, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### I. D. d) Try some different groupings, such as 6, 6, 6, 6, 3, 6, 6, 6, 6 (and no median.row)
mmplot(stat.data=edPov,
       map.data=statePolys,
       map.link=c('StateAb', 'ID'),
       panel.types=c('map', 'dot_legend', 'labels', 'dot', 'dot'),
       panel.data=list(NA, NA, 'state', 'ed', 'pov'),
       ord.by='ed',   
       rev.ord=TRUE,
       grouping=c(6, 6, 6, 6, 3, 6, 6, 6, 6), median.row=FALSE)
```



```{r Ch2-micromap9, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### I. D. e) Align the state names to the left, see panel.att page 5-6 User guide
mmplot(stat.data=edPov,
       map.data=statePolys,
       map.link=c('StateAb', 'ID'),
       panel.types=c('map', 'dot_legend', 'labels', 'dot', 'dot'),
       panel.data=list(NA, NA, 'state', 'ed', 'pov'),
       ord.by='ed',   
       rev.ord=TRUE,
       grouping=c(6, 6, 6, 6, 3, 6, 6, 6, 6), median.row=FALSE,
       panel.att=list(list(3, align='left')))
```



```{r Ch2-micromap10, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### I. D. f) Vertically align the state names to the center, see vertical.align page 17 User guide
mmplot(stat.data=edPov,
       map.data=statePolys,
       map.link=c('StateAb', 'ID'),
       panel.types=c('map', 'dot_legend', 'labels', 'dot', 'dot'),
       panel.data=list(NA, NA, 'state', 'ed', 'pov'),
       ord.by='ed',   
       rev.ord=TRUE,
       grouping=c(6, 6, 6, 6, 3, 6, 6, 6, 6), median.row=FALSE,
       vertical.align='center',
       panel.att=list(list(3, align='left')))
```



```{r Ch2-micromap11, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### I. D. g) Use a divergent 6-class BrBG color scheme from RColorBrewer (which is colorblind safe)
mmplot(stat.data=edPov,
       map.data=statePolys,
       map.link=c('StateAb', 'ID'),
       panel.types=c('map', 'dot_legend', 'labels', 'dot', 'dot'),
       panel.data=list(NA, NA, 'state', 'ed', 'pov'),
       ord.by='ed',   
       rev.ord=TRUE,
       grouping=c(6, 6, 6, 6, 3, 6, 6, 6, 6), median.row=FALSE,
       vertical.align='center',
       colors=brewer.pal(6, "BrBG"),
       panel.att=list(list(3, align='left')))
```


```{r Ch2-micromap12, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### I. E. Refine the Micromap Plot
mmplot(stat.data=edPov, map.data=statePolys,
        panel.types=c('dot_legend',  'labels', 'dot', 'dot', 'map'),
        panel.data=list(NA, 'state', 'pov', 'ed', NA),
        map.link=c('StateAb', 'ID'),
        ord.by='pov', 
        grouping=5, 
        median.row=TRUE, 
        
        plot.height=9, 
        
        colors=c('red', 'orange', 'green', 'blue', 'purple'),
        
        panel.att=list(list(1, point.type=20, point.border=TRUE, point.size=2),
                       
                       list(2, header='States', panel.width=.8, 
                            align='left', text.size=.9),
                       
                       list(3, header='Percent Living Below\nPoverty Level',
                            graph.bgcolor='lightgray', point.size=1.5,
                            xaxis.ticks=list(10, 15, 20), xaxis.labels=list(10, 15, 20),
                            xaxis.title='Percent'),
                       
                       list(4, header='Percent Adults With\n4+ Years of College',
                            graph.bgcolor='lightgray', point.size=1.5,
                            xaxis.ticks=list(20, 30, 40), xaxis.labels=list(20, 30, 40), 
                            xaxis.title='Percent'),
                       
                       list(5, header='Light Gray Means\nPreviously Displayed',
                            map.all=TRUE, fill.regions='aggregate',
                            active.border.color='black', active.border.size=1.5,
                            inactive.border.color=gray(.7), inactive.border.size=1, 
                            panel.width=.8)))
```



```{r Ch2-micromap13, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### I. F. Refine the Micromap Plot with adjusted ticmarks & ticmark labels
mmplot(stat.data=edPov, map.data=statePolys,
       panel.types=c('dot_legend',  'labels', 'dot', 'dot', 'map'),
       panel.data=list(NA, 'state', 'pov', 'ed', NA),
       map.link=c('StateAb', 'ID'),
       ord.by='pov', 
       grouping=5, 
       median.row=TRUE, 
       
       plot.height=9, 
       
       colors=c('red', 'orange', 'green', 'blue', 'purple'),
       
       panel.att=list(list(1, point.type=20, point.border=TRUE, point.size=2),
                      
                      list(2, header='States', panel.width=.8, 
                           align='left', text.size=.9),
                      
                      list(3, header='Percent Living Below\nPoverty Level',
                           graph.bgcolor='lightgray', point.size=1.5,
                           xaxis.ticks=as.list(extended(min(edPov$pov), max(edPov$pov), 3)), 
                           xaxis.labels=as.list(extended(min(edPov$pov), max(edPov$pov), 3)),
                           xaxis.title='Percent'),
                      
                      list(4, header='Percent Adults With\n4+ Years of College',
                           graph.bgcolor='lightgray', point.size=1.5,
                           xaxis.ticks=as.list(extended(min(edPov$ed), max(edPov$ed), 4)), 
                           xaxis.labels=as.list(extended(min(edPov$ed), max(edPov$ed), 4)), 
                           xaxis.title='Percent'),
                      
                      list(5, header='Light Gray Means\nPreviously Displayed',
                           map.all=TRUE, fill.regions='aggregate',
                           active.border.color='black', active.border.size=1.5,
                           inactive.border.color=gray(.7), inactive.border.size=1, 
                           panel.width=.8)))
```



```{r Ch2-micromap14, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
hist(1:100)
```


## Code Chunks China {#Ch2-CodeChunksChina}


```{r Ch2-micromaChina1, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
#VI. Use of external shapefiles

library(rgeos)
library(rgdal)
library(maptools)

library(micromap)

### Read in the shapefile and simplify the polygons (run this part of the code only once!)

ChinaShapefile <- readOGR("data/China_Shapefiles", "export", verbose = TRUE)
plot(ChinaShapefile)
summary(ChinaShapefile)
```


```{r Ch2-micromaChina2, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
gIsValid(ChinaShapefile, byid = TRUE, reason = TRUE)
ChinaShapefileThin <- thinnedSpatialPoly(ChinaShapefile, 
                                         tolerance = 300, # try 100, 300, 3000, 30000
                                         minarea = 10000000000,
                                         topologyPreserve = TRUE, 
                                         avoidGEOS = TRUE)
ChinaShapefileThin <- gBuffer(ChinaShapefileThin, width = 0, byid = TRUE)
gIsValid(ChinaShapefileThin, byid = TRUE, reason = TRUE)
plot(ChinaShapefileThin)

### Change tolerance to 100, 3000, and 30000 and run the the entire code again, starting with readOGR.
### Which of these is the best choice for tolerance? You may have to enlarge the map.

### If you plan to use LM plots for a research project, you may also want to 
### enlarge small regions or shift far-away regions closer to the main region.
### Talk to me for further information.
```


```{r Ch2-micromaChina3, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### Read in the data

religion <- read.csv("data/China_ReligionData.csv", header = TRUE)

ChinaPolys <- create_map_table(ChinaShapefileThin, "ename")

### Basic micromap plot

ChinaPlot <- mmplot(stat.data = religion,
                    map.data = ChinaPolys,
                    panel.types = c("labels", "dot", "map"),
                    panel.data = list("Province", "Christianity", NA),
                    ord.by = "Christianity",
                    grouping = 5,
                    median.row = TRUE, 
                    map.link = c("Province", "ID"))
```


```{r Ch2-micromaChina4, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### Advanced micromap plot

ChinaReligion <- mmplot(stat.data = religion, 
                        map.data = ChinaPolys,
                        panel.types = c("map", "dot_legend", "labels", 
                                        "dot", "dot", "dot", "dot"),
                        panel.data = list(NA, NA, "Province", 
                                          "Christianity", "Buddhism", "Daoism", "Islam"),
                        map.link = c("Province", "ID"),
                        ord.by = "Christianity",
                        rev.ord = TRUE,
                        
                        grouping = 5,
                        median.row = TRUE,
                        plot.height = 9,
                        plot.width = 9,
                        colors = c("red", "orange", "green", "blue", "purple"),
                        two.ended.maps = TRUE,
                        map.all = TRUE,
                        map.color2 = "lightgray",
                        
                        plot.header = "Religion in China",
                        plot.header.size = 2,
                        plot.header.color = "black",
                        
                        plot.panel.spacing = 0,
                        panel.att = list(
                          list(1, header = "Two-ended\nCumulative Maps",
                               inactive.border.color = gray(.7), 
                               inactive.border.size = 1,
                               panel.width = 1.2),
                          list(2, point.type = 20,
                               point.size = 1.4),
                          list(3, header = "Provinces", 
                               panel.width = .9,
                               align = "left", 
                               text.size = .8),
                          list(4, header = "Christianity",
                               header.color = "red",
                               graph.bgcolor = "lightgray", 
                               point.size = 1,
                               xaxis.ticks = seq(0, 3000, by = 1000), 
                               xaxis.labels = seq(0, 3, by = 1),
                               xaxis.title = "Number (Thousand)"
                          ),
                          list(5, header = "Buddhism",
                               graph.bgcolor = "lightgray", 
                               point.size = 1,
                               xaxis.ticks = seq(0, 3000, by = 1000), 
                               xaxis.labels = seq(0, 3, by = 1),
                               xaxis.title = "Number (Thousand)"
                          ),
                          list(6, header = "Daoism",
                               graph.bgcolor = "lightgray", 
                               point.size = 1,
                               xaxis.ticks = seq(0, 2000, by = 1000), 
                               xaxis.labels = seq(0, 2, by = 1),
                               xaxis.title = "Number (Thousand)"
                          ),
                          list(7, header = "Islam",
                               graph.bgcolor = "lightgray", 
                               point.size = 1,
                               xaxis.ticks = seq(0, 25000, by = 12500), 
                               xaxis.labels = seq(0, 25, by = 12.5),
                               xaxis.title = "Number (Thousand)"
                          )
                        )
)
```


```{r Ch2-micromaChina5, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### Final advanced micromap plot with reduced margins

ChinaReligion <- mmplot(stat.data = religion, 
                        map.data = ChinaPolys,
                        panel.types = c("map", "dot_legend", "labels", 
                                        "dot", "dot", "dot", "dot"),
                        panel.data = list(NA, NA, "Province", 
                                          "Christianity", "Buddhism", "Daoism", "Islam"),
                        map.link = c("Province", "ID"),
                        ord.by = "Christianity",
                        rev.ord = TRUE,
                        
                        grouping = 5,
                        median.row = TRUE,
                        plot.height = 9,
                        plot.width = 9,
                        colors = c("red", "orange", "green", "blue", "purple"),
                        two.ended.maps = TRUE,
                        map.all = TRUE,
                        map.color2 = "lightgray",
                        
                        plot.header = "Religion in China",
                        plot.header.size = 2,
                        plot.header.color = "black",
                        
                        plot.panel.spacing = 0,
                        panel.att = list(
                          list(1, header = "Two-ended\nCumulative Maps",
                               inactive.border.color = gray(.7), 
                               inactive.border.size = 1,
                               panel.width = 1.2),
                          list(2, point.type = 20,
                               point.size = 1.4),
                          list(3, header = "Provinces", 
                               panel.width = .9,
                               align = "left", 
                               text.size = .8),
                          list(4, header = "Christianity",
                               header.color = "red",
                               graph.bgcolor = "lightgray", 
                               point.size = 1,
                               right.margin = 0, 
                               left.margin = -0.5,
                               xaxis.ticks = seq(0, 3000, by = 1000), 
                               xaxis.labels = seq(0, 3, by = 1),
                               xaxis.title = "Number (Thousand)"
                          ),
                          list(5, header = "Buddhism",
                               graph.bgcolor = "lightgray", 
                               point.size = 1,
                               right.margin = 0, 
                               left.margin = -0.5,
                               xaxis.ticks = seq(0, 3000, by = 1000), 
                               xaxis.labels = seq(0, 3, by = 1),
                               xaxis.title = "Number (Thousand)"
                          ),
                          list(6, header = "Daoism",
                               graph.bgcolor = "lightgray", 
                               point.size = 1,
                               right.margin = 0, 
                               left.margin = -0.5,
                               xaxis.ticks = seq(0, 2000, by = 1000), 
                               xaxis.labels = seq(0, 2, by = 1),
                               xaxis.title = "Number (Thousand)"
                          ),
                          list(7, header = "Islam",
                               graph.bgcolor = "lightgray", 
                               point.size = 1,
                               right.margin = 0.25, 
                               left.margin = -0.5,
                               xaxis.ticks = seq(0, 25000, by = 12500), 
                               xaxis.labels = seq(0, 25, by = 12.5),
                               xaxis.title = "Number (Thousand)"
                          )
                        )
)
```


## Further Reading {#Ch2-FurtherReading}


Introduce cross-references to other chapters, e.g., Chapter \@ref(Ch1) and Chapter \@ref(Ch3),
where related work and further examples can be found in this book that match the content of this
chapter, that follow up on this chapter, or that are a prerequisite of this chapter.

Also, do some scientific literature review here that is specific to your chapter.
Where has this R package been introduced and used before, where have other plot types
or different countries been used in micromaps, what were other applications 
of micromaps that are related to the title and content of your chapter, etc.?


\printbibliography[segment=\therefsegment,heading=subbibliography]

