# Linked Micromap Plots via the **micromap** R Package {#Ch2}


\chapterauthor{J{\"u}rgen Symanzik, Marcus W. Beck, Michael G. McManus}


The **micromap**\index{R Packages!micromap} R package [@PaOl2015],
accessible at https://cran.r-project.org/web/packages/micromap/index.html,
will be introduced in this chapter. The reader will learn how to create a 
basic linked micromap plot\index{Linked micromap plot} via this R package. 
Details will be provided how to optimize and fine-tune such a basic plot 
into a publication-worthy final linked micromap plot.


## Introduction {#Ch2-Introduction}


Write chapter introduction here.


## Chapter Outline (Merge into Chapter Introduction at the End) {#Ch2-Outline}


Note: Chapter \@ref(Ch1) will cover elements of a linked micromaps, 4 panels of legend, label, statistic, micromaps 
and rows of perceptual groups? Also discussion of spatial patterns and interpretation of positive, negative
(or no) association in statistical panels of dot plots.\index{Dot plot}


-	Overview of Four Steps (**Marcus**: Maybe a good figure could be a conceptual diagram showing the whole process?)
    - Identifying and Geoprocessing of Spatial Boundary Data
    - Linking Spatial Boundary Data and Statistical Data
    - Creating a Draft Linked Micromap Plot
    - Refining the Linked Micromap Plot
    
-	Spatial Polygons in R and micromap Package Updates (**Juergen**: How much of this do we need here in Chapter \@ref(Ch2)? Would it make more sense to mention this in Chapter \@ref(Ch4)?)
    - Spatial objects and simple feature objects
    - Updates
    
-	Identifying and Geoprocessing of Spatial Boundary Data (**Marcus**: I think this is a good topic to discuss, but may be tricky to balance specificity with brevity.  I'm no expert on it, but I know there are a few different methods one could use.  Maybe just hit on the why and how, and briefly present tradeoffs of the different methods?  I think for MM, the goal is just to increase render time w/o sacrificing spatial specificity, so maybe the exact method is not important? -- **Juergen**: In fact, let's keep it simple here and work with ready-to-use shapefiles that are already available in some R packages; refer to Chapter \@ref(Ch4) how to bring in external shapefiles and modify them for use in micromaps)
    - Keep this brief as Chapter \@ref(Ch4) might cover this in detail?
    - Example of generalizing polygons Figure  (Use same figures as Chapter \@ref(Ch4)?)

-	Linking Spatial Boundary Data and Statistical Data
    - Explicit link to spatial object
    - Under the hood link to simple features object

-	Creating a Draft Linked Micromap Plot
    - Use workshop NARS pH sample & population estimates Figure (**Juergen**: Use simpler edPov data in first example)

-	Refining the Linked Micromap Plot
    - Use workshop NARS pH sample & population estimates Figure (**Juergen**: Use simpler edPov data in first example)

- **Juergen**: Introducing a second basic example for a different geographic region, e.g., state level or environmental, and use of a different built-in plot type (e.g., boxplot). Similar four steps as before, but more condensed.

-	Challenges to Visualizing Data with Linked Micromap Plots (**Marcus**: Not sure this is the right place for it, but maybe it's worth providing a brief contrast with micromapST?  Maybe include this in the overview at the top?  I think it would be useful for the reader to understand succinctly how the two differ and why there are two packages, though I'm sure these themes will be present throughout the book.  Perhaps the latter point is not that interesting, i.e., it's common to have "competing" software in an open source environment? -- **Juergen**: Yes, we need some discussions at the end of Chapters \@ref(Ch2) and \@ref(Ch3) that outline advantages and limitations of each of the two main micromap packages. Alternatively, there could be some table in Chapter \@ref(Ch1) that lists features (and limitations) of the two packages when they are first introduced to the reader. Overall, this last point may be a subsection in the Introduction, or become a final chapter at the end of the book which research is necessary in the future.)
    - Problem of many polygons (**Juergen**: There are other problems, e.g., many islands as in the Philippines, narrow shapes as for Chile, subregions of considerably different sizes as for Russia)
    - Automating linked micromap plot production for reports and interactive visualizations (Check with Emma Jones at VA DEQ)
    - (**Marcus**: Related, a disadvantage of both packages is their lack of tighter integration with ggplot.  Not sure we want to go there though. -- **Juergen**: I wouldn't go there as these are the packages that currently exist and have to be used.)


## Steps to Create a Linked Micromap Plot with the **micromap**\index{R Packages!micromap} R Package {#Ch2-Steps}


Four main steps, shown in Figure \@ref(fig:Ch2-flowchart), are needed to create a 
linked micromap plot\index{Linked micromap plot} 
with the **micromap**\index{R Packages!micromap} R package:

1. **Identifying and Geoprocessing of Spatial Boundary Data** (see Section \@ref(Ch2-Identifying) for details): 
In addition to a data frame that contains the statistical data,
the user must identify a data structure that contains the spatial boundary data for the region and subregions that have
to be colored in the maps. This spatial boundary data typically comes in the form of a
SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame}. In this chapter, we will work with ready-to-use
boundary files. In Chapter \@ref(Ch4), we will discuss how to make use of boundary files that are
provided as external shapefiles\index{Shapefiles}. We will also see in Chapter \@ref(Ch4) how
to simplify complex boundaries, enlarge small subregions in the maps, and move subregions that are far away closer
to the main area of the map.

2. **Linking Spatial Boundary Data and Statistical Data** (see Section \@ref(Ch2-Linking) for details): 
To link spatial boundary data and statistical data, 
the SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame} first has to be transformed into a regular data frame
via the `micromap::create_map_table()` function.
Next, we have to identify one variable from the statistical data frame
and one variable from the newly created data frame with the boundary information that 
allow us to link statistical data and boundary data for each subregion.

3. **Creating a Draft Linked Micromap Plot** (see Section \@ref(Ch2-Creating) for details): 
While not necessary, it is always a good idea to first create a minimal
linked micromap plot\index{Linked micromap plot} to ensure that the statistical data and boundary data
are matching and a correct draft linked micromap plot\index{Linked micromap plot} is created. Skipping this
step and trying to create a complex linked micromap plot\index{Linked micromap plot} immediately may make it hard
to find possible errors in the R code.

4. **Refining the Linked Micromap Plot** (see Section \@ref(Ch2-Refining) for details): 
Once a draft linked micromap plot\index{Linked micromap plot} has been
created, this plot usually needs fine-tuning of its appearance and plot aesthetics, e.g., modification of colors,
change of the layout and of perceptual groups, addition of labels and legends, and possibly the addition of additional
statistical variables or changes to different graph types for some of the variables.


```{r Ch2-flowchart, fig.cap = 'Workflow to create a linked micromap plot with the **micromap**\\index{R Packages!micromap} R package (Diagram created with the **DiagrammeR**\\index{R Packages!DiagrammeR} R package [@Iannone2022]).', fig.width = 5, fig.height = 5, echo = FALSE}
library(DiagrammeR)

grViz(
  diagram = "digraph flowchart {
  node [fontname = arial, shape = oval, color = grey, style = filled]
  tab1 [label = '@@1']
  tab2 [label = '@@2']
  tab3 [label = '@@3']
  tab4 [label = '@@4']

  tab1 -> tab2 -> tab3 -> tab4;
}

  [1]: '1. Identifying and Geoprocessing of Spatial Boundary Data'
  [2]: '2. Linking Spatial Boundary Data and Statistical Data'
  [3]: '3. Creating a Draft Linked Micromap Plot'
  [4]: '4. Refining the Linked Micromap Plot'
"
)
```


## Example 1: A Linked Micromap Plot for the 50 US States and Washington, D.C. {#Ch2-Example1}


In this first linked micromap plot\index{Linked micromap plot}
example created with the **micromap**\index{R Packages!micromap} R package, we work with the
_USstates_\index{Datasets!USstates} and _edPov_\index{Datasets!edPov} datasets
from the **micromap**\index{R Packages!micromap} R package.
We follow the four steps outlined in Section \@ref(Ch2-Steps).


### Identifying and Geoprocessing of Spatial Boundary Data {#Ch2-Identifying}


_USstates_\index{Datasets!USstates} is a SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame} for 
the 50 states (and Washington, D.C.) of the United States (U.S.) 
that was created for use in linked micromap plots\index{Linked micromap plot}. Notably, the boundaries
of many of the states have been simplified, Alaska and Hawaii have been moved closer to the
contiguous 48 states (and also have been resized), and Washington, D.C., has been pulled out of the
main map and has been placed further to the east (and also has been enlarged)
as shown in Figure \@ref(fig:Ch2-USstates). In the following R code,
we first load the **micromap**\index{R Packages!micromap} R package and the
_USstates_\index{Datasets!USstates} data set, verify that this object indeed is a
SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame}, then look
at some of the data in the `data` component of this object, and finally plot it.


```{r Ch2-USstates, fig.cap = 'Map representation of the _USstates_\\index{Datasets!USstates} spatial boundary data set for the United States that frequently is used as the basis for linked micromap plots\\index{Linked micromap plot} that are created with the **micromap**\\index{R Packages!micromap} R package.', fig.width = 7, fig.height = 4}
library(micromap)

data(USstates)

class(USstates)
head(USstates@data)

plot(USstates)
```


### Linking Spatial Boundary Data and Statistical Data {#Ch2-Linking}


In this step, the SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame} first is transformed
into a regular data frame for use in a linked micromap plot\index{Linked micromap plot}
via the `micromap::create_map_table()` function. We have to indicate a variable from the `data`
component of the SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame} object that 
can be used as an ID column. 
A matching ID column has to be identified in the statistical dataset. These two ID columns
have to be used for linking the two data sets. For the _USstates_\index{Datasets!USstates} dataset,
this is usually the `ST` variable that contains the 50 abbreviations for the US states (and for D.C.).
The resulting `state_polys_table` is a regular data frame that can be used for creating a
linked micromap plot\index{Linked micromap plot} in the next step.
Note that there are typically more rows in this data frame than there are spatial subregions.
This is because many spatial subregions consist of more than just one boundary, e.g., in the
presence of islands or if a land mass is split into more than one area, e.g., as in Michigan in the US.


```{r Ch2-linkingdataframesUSstates}
head(USstates@data$ST)

state_polys_table <- create_map_table(
  tmp.map = USstates,
  IDcolumn = "ST"
)

class(state_polys_table)
dim(state_polys_table)
names(state_polys_table)
```


_edPov_\index{Datasets!edPov} is a data frame that contains
education and poverty level data for the 50 states (and Washington, D.C.).
This data frame has 51 rows, one for each of the 50 states (and for D.C.).
We have to identify one variable from this data frame that can be used for linking the
two data sets. Here, this is the `StateAb` variable.

The last expression in the following R code verifies that there is indeed at least one matching ID 
in the `state_polys_table` data frame for each ID in the _edPov_\index{Datasets!edPov} dataset.
Mismatching or missing IDs in the statistical data frame may cause
that data for some subregions do not appear
on any of the maps later on or that some subregions shown on the maps do not get colored in any of the maps.
While this often can be contributed to missing data in the statistical data frame, e.g., there is
only data for the 50 states but not for Washington, D.C., mismatching identifiers also can be
the underlying reason (e.g., if the statistical data frame uses `D.C.` as ID while
the spatial data frame uses `DC`). Here, everything is matching.


```{r Ch2-linkingdataframesedPov}
data(edPov)
dim(edPov)

head(edPov)
head(edPov$StateAb)

all(sort(edPov$StateAb) == sort(unique(state_polys_table$ID)))
```


### Creating a Draft Linked Micromap Plot {#Ch2-Creating}


Now we can create a minimal draft linked micromap plot\index{Linked micromap plot} using the
`micromap::lmplot()` function, based on the previously created 
`state_polys_table` data frame and the _edPov_\index{Datasets!edPov} dataset.
In our function call, we only provide seven required arguments. None of these arguments has a default setting.
For all other arguments of this function, the default settings will be used here. 
The statistical data (_edPov_\index{Datasets!edPov}) is assigned to the `stat.data` argument
and the spatial boundary data (in the `state_polys_table` data frame) is assigned to the `map.data` argument.

The `map.link` argument is needed to tell the function how to link the statistical data
and the spatial boundary data. A vector with the names of the two linking variables 
identified in the previous step is needed for this argument.
The first variable name (`StateAb`) must come from the statistical data (here _edPov_\index{Datasets!edPov})
and the second variable name (`ID`) must come from the `state_polys_table` data frame that
contains the spatial boundary information.
Changing the order of these two variable names typically results in an error.

The `panel.types` and `panel.data` are closely related. As the name suggests, `panel.types` is a vector
that specifies the layout of the columns in the linked micromap plot\index{Linked micromap plot}.
Here, we have the `dot_legend` in the first column, `labels` in the second column,
two columns with statistical data represented as dot plots\index{Dot plot} (`dot`),
and the micromaps in the fifth and final (`map`) column.
This is matched with a list of data that is used for each of the five columns.
A list is necessary for this argument as some of the data itself can be lists as we will
see in the R code for Figures \@ref(fig:Ch2-creatingdraftWV) 
and \@ref(fig:Ch2-refining1WV) later on.

The `dot_legend` simply shows a plotting symbol in a certain color that represents that row in
the linked micromap plot\index{Linked micromap plot}. Thus, no further data is needed and `NA`
is assigned. `labels` requires some text argument, typically some identifiers of the subregions
in the maps. Here `state` from _edPov_ is used. 
The variables `pov` and `ed` from _edPov_ are used for the statistical displays in columns three and four.
It should be noted that all data specified in `panel.data` by default is taken from the
data frame specified in the `stat.data` argument.
The fifth and final column contains the micromaps. Their boundaries are obtained from
the `map.data` data frame. Thus, no further data has to be specified here. Instead, `NA` 
is assigned here.

Two more arguments have to be specified: `ord.by` specifies the sorting variable of the 
rows in the linked micromap plot\index{Linked micromap plot}. Here, `pov` 
from the `stat.data` argument is used. The sorting of the rows goes from smallest (on top) to
largest (at the bottom). Finally, `grouping` specifies  the number of rows
in each of the perceptual groups (here `5` rows). If a single integer value is provided, 
that value is used for all perceptual groups. If a vector of integer values is provided,
each perceptual group may have a different number of rows as 
shown in the R code for Figure \@ref(fig:Ch2-refining3).

The resulting draft linked micromap plot\index{Linked micromap plot} is shown in
Figure \@ref(fig:Ch2-creatingdraft). It is noteworthy that there are eleven perceptual groups
overall --- ten with five subregions and one with just one subregion. This is due to the fact
that there are 51 subregions overall: The 50 US states and Washington, D.C.
We abstain from interpreting this figure at this stage, but we will provide some
helpful interpretation once we have created the last refined version of this figure
in the next step.


```{r Ch2-creatingdraft, fig.cap = 'Draft linked micromap plot\\index{Linked micromap plot}, based on the _edPov_\\index{Datasets!edPov} dataset.', fig.width = 7, fig.height = 9}
lmplot(
  stat.data = edPov,
  map.data = state_polys_table,
  map.link = c("StateAb", "ID"),
  panel.types = c("dot_legend", "labels", "dot", "dot", "map"),
  panel.data = list(NA, "state", "pov", "ed", NA),
  ord.by = "pov",
  grouping = 5
)
```


### Refining the Linked Micromap Plot {#Ch2-Refining}


We continue with the draft linked micromap plot\index{Linked micromap plot} from the previous
step and refine it in multiple small steps. This refinement should only be started once
functional R code has been obtained in the previous step.

First, we get rid of the eleventh perceptual group (with just one subregion) and introduce
a median row via `median.row = TRUE` instead. A median row is often a good solution if the 
number of subregions is odd such as for the 51 states (and D.C.) here.
Here, Wyoming is the state shown in the median row. It has the 
26$^{th}$ highest (or lowest) value for the sorting variable, i.e., `pov`.
It does not appear in a map by itself, but rather is added to the perceptual
groups above and below the median row in a neutral color, 
thus increasing the number of subregions shown in each of these two maps by one (i.e., six here)
Also, we reverse the sorting order via `rev.ord = TRUE`. 
Now the sorting of the rows goes from largest (on top) to
smallest (at the bottom).
The resulting linked micromap plot\index{Linked micromap plot} is shown in
Figure \@ref(fig:Ch2-refining1).
While we keep `pov` as the sorting variable in this first refined version,
the reader is encouraged to use `ed` as sorting variables and see how
the spatial patterns highlighted in the maps change. This can be done
for the original sorting order or for the reversed sorting order.
The rows can even be sorted by `region` or alphabetically 
by `state` or `StateAb` even though such an alphabetical sorting
in most cases is not very meaningful, at least not for this specific example.


```{r Ch2-refining1, fig.cap = 'First refined linked micromap plot\\index{Linked micromap plot}, based on the _edPov_\\index{Datasets!edPov} dataset. Main changes are the introduction of a median row instead of the eleventh preceptual group and the reverse ordering of the `pov` data in the first statistical column.', fig.width = 7, fig.height = 9}
lmplot(
  stat.data = edPov,
  map.data = state_polys_table,
  map.link = c("StateAb", "ID"),
  panel.types = c("dot_legend", "labels", "dot", "dot", "map"),
  panel.data = list(NA, "state", "pov", "ed", NA),
  ord.by = "pov",
  rev.ord = TRUE,
  grouping = 5,
  median.row = TRUE
)
```


We continue with this first refined linked micromap plot\index{Linked micromap plot}
and make further modifications. Next, we change the order of the two statistical columns,
i.e., we place `ed` to the left of `pov` and use `ed` as the sorting variable (in reverse order).
While any column of the linked micromap plot\index{Linked micromap plot} or even
variables not shown in it can be used as sorting variables, in most cases,
the first (leftmost) statistical column is used for sorting purposes.
Moreover, we place the maps on the left side of the plot. As previously stated,
the `panel.types` and `panel.data` arguments of the `micromap::lmplot()` function
are closely related. Thus, if we change the order of one, the order of the
other one has to be changed accordingly. 
The resulting linked micromap plot\index{Linked micromap plot} is shown in
Figure \@ref(fig:Ch2-refining2).
There is no strong recommendation where the column with the maps should be placed.
This often depends on personal preferences of the authors.
There exist examples in the literature where the maps are placed on the left,
e.g., in @Mast2013LinkedMicromaps and @DaWo2021, 
on the right, 
e.g., in @Wartenberg2009,
and even in the middle,
e.g., in @TaSt2019.


```{r Ch2-refining2, fig.cap = 'Second refined linked micromap plot\\index{Linked micromap plot}, based on the _edPov_\\index{Datasets!edPov} dataset. Main changes are related to the order of the five columns in the plot. Most notable, the map column is shown on the left here.', fig.width = 7, fig.height = 9}
lmplot(
  stat.data = edPov,
  map.data = state_polys_table,
  map.link = c("StateAb", "ID"),
  panel.types = c("map", "dot_legend", "labels", "dot", "dot"),
  panel.data = list(NA, NA, "state", "ed", "pov"),
  ord.by = "ed",
  rev.ord = TRUE,
  grouping = 5,
  median.row = TRUE
)
```


We continue making changes to the second refined linked micromap plot\index{Linked micromap plot}.
We change the grouping to nine perceptual groups overall (and no median row)
via `grouping = c(6, 6, 6, 6, 3, 6, 6, 6, 6)` and `median.row = FALSE`
and vertically align the rows in each perceptual block via `vertical.align = "center"`.
Finally, we start making changes to individual columns of the plot via
the `panel.att` argument. Here, the third column that shows the `labels` is aligned on the left.
The resulting linked micromap plot\index{Linked micromap plot} is shown in
Figure \@ref(fig:Ch2-refining3).


```{r Ch2-refining3, fig.cap = 'Third refined linked micromap plot\\index{Linked micromap plot}, based on the _edPov_\\index{Datasets!edPov} dataset. Main changes are related to perceptual groups with different numbers of subregions and the vertical alignment of rows in the middle of the plot. Also, `labels` are aligned on the left.', fig.width = 7, fig.height = 9}
lmplot(
  stat.data = edPov,
  map.data = state_polys_table,
  map.link = c("StateAb", "ID"),
  panel.types = c("map", "dot_legend", "labels", "dot", "dot"),
  panel.data = list(NA, NA, "state", "ed", "pov"),
  ord.by = "ed",
  rev.ord = TRUE,
  grouping = c(6, 6, 6, 6, 3, 6, 6, 6, 6),
  median.row = FALSE,
  vertical.align = "center",
  panel.att = list(list(3, align = "left"))
)
```


After this experiment with a different grouping, we revert back to the more traditional grouping
of ten perceptual groups with five rows each and a median row that is frequently used
for the 50 US states (and D.C.). What we want to discuss next is the choice of colors
for the subregions in each map (and thus for the `dot_legend` appearance in the
`panel.type` argument as well). The default setting for the `colors` argument
makes use of `max(grouping)` different colors from a spectral color scheme.
Thus, when `grouping = 5` as in Figures \@ref(fig:Ch2-creatingdraft)-\@ref(fig:Ch2-refining2),
a five-class spectral color scheme\index{Color scheme!Spectral} is selected, 
whereas a six-class spectral color scheme\index{Color scheme!Spectral} is
selected when `grouping = c(6, 6, 6, 6, 3, 6, 6, 6, 6)` as in Figure \@ref(fig:Ch2-refining3).
Historically, many linked micromap plots,\index{Linked micromap plot} 
e.g., in @COCPC1998 and @WCCBP2002,
made use of rainbow colors\index{Color scheme!Rainbow colors} that could be obtained 
via the `colors = c("red", "orange", "green", "blue", "purple")` setting of the
`colors` argument. While these colors work well for readers with normal color vision,
they may not work well for readers with certain types of color vision deficiencies.
Instead, some color schemes that are colorblind safe are better suited for such readers.
Options are single-hue or multi-hue sequential color schemes\index{Color scheme!Sequential} 
or selected divergent color schemes\index{Color scheme!Divergent}.
Such color schemes can be obtained from the 
**RColorBrewer**\index{R Packages!RColorBrewer} R package [@Neuwirth2022].
The reader is encouraged to read more about the theoretical background
of these color schemes in @BHH2003 and @HaBr2003 and experiment with different settings
at the supporting web page at https://colorbrewer2.org/.
@SDWPM2014 used a five-class greyscale sequential color scheme\index{Color scheme!Sequential}
from **RColorBrewer**\index{R Packages!RColorBrewer} in reverse sorting (where
the darkest grey color comes first) for publication in a greyscale publication,\index{Colors!Greyscale publication}
obtained via `colors = brewer.pal(n = 5, name = "Greys")[5:1]`.
@SBDSS2016 used a five-class divergent red-yellow-blue (RdYlBu) color scheme\index{Color scheme!Divergent}
from **RColorBrewer**\index{R Packages!RColorBrewer}
that is colorblind safe\index{Colors!Colorblind safe} and print friendly,\index{Colors!Print friendly}
obtained via `colors = brewer.pal(n = 5, name = "RdYlBu")`.
In the linked micromap plot\index{Linked micromap plot} shown in
Figure \@ref(fig:Ch2-refining4), we use a
five-class divergent brown-blue-green (BrBG) color scheme\index{Color scheme!Divergent}
from **RColorBrewer**\index{R Packages!RColorBrewer}
that is also colorblind safe\index{Colors!Colorblind safe} and print friendly,\index{Colors!Print friendly}
obtained via `colors = brewer.pal(n = 5, name = "BrBG")`.


```{r Ch2-refining4, fig.cap = 'Fourth refined linked micromap plot\\index{Linked micromap plot}, based on the _edPov_\\index{Datasets!edPov} dataset. Main changes are the use of a divergent brown-blue-green color scheme\\index{Color scheme!Divergent} and the conversion back to the traditional grouping for 51 subregions.', fig.width = 7, fig.height = 9}
lmplot(
  stat.data = edPov,
  map.data = state_polys_table,
  map.link = c("StateAb", "ID"),
  panel.types = c("map", "dot_legend", "labels", "dot", "dot"),
  panel.data = list(NA, NA, "state", "ed", "pov"),
  ord.by = "ed",
  rev.ord = TRUE,
  grouping = 5,
  median.row = TRUE,
  colors = brewer.pal(n = 5, name = "BrBG"),
  panel.att = list(list(3, align = "left"))
)
```


So far, there are no labels for the columns and no titles for the statistical columns shown.
Tic marks and tic mark labels, in particular in the second statistical column, could be improved,
background colors in the statistical columns and maps could be modified, 
font and symbol sizes could be modified, and the widths of the columns could be adjusted.
All of this is done via the `panel.att` argument that controls the panel specific attributes of
each column in the linked micromap plot.\index{Linked micromap plot}
The content of this argument typically is a list of lists where the attributes
for each column of the linked micromap plot\index{Linked micromap plot} are modified via a separate list.
These inner lists are numbered from 1 to the number of elements in the `panel.types` vector
where `1` is related to `map`, `2` to `dot_legend`, `3` to `labels`, and `4` and `5` to `dot`, i.e.,
the dot plots\index{Dot plot} in the two statistical columns,
in the linked micromap plot\index{Linked micromap plot} shown in
Figure \@ref(fig:Ch2-refining5).

We leave it to the reader to further experiment with the different elements in these lists.
If the purpose of a certain element is not immediately obvious, it might be a good idea
to considerably increase or decrease the numeric value of that element or change the color to
`red` or `yellow` to highlight that element. 

Here, we only want to explain the purpose of the `fill.regions` element with the matching `header` element
in the list for column `1`, i.e., the `map` column:
The setting `fill.regions = "aggregate"` (which in fact is the default setting)
fills in the subregions from all previous perceptual groups in the
subsequent perceptual groups. This filing proceeds from
the top perceptual group to the bottom perceptual group by sequentially
filling the subregions that have already been displayed.
Thus, in the map for the final perceptual group at the bottom, all subregions have been filled.
The text in the `header` is used to communicate this information to the reader.
Alternatively, the setting `fill.regions = "with data"` only fills those subregions
in a map that actually show data in that perceptual group. 
No additional subregions are filled in any of the maps.
Another setting for `fill.regions` is discussed in the next refinement step.


```{r Ch2-refining5, fig.cap = 'Fifth refined linked micromap plot\\index{Linked micromap plot}, based on the _edPov_\\index{Datasets!edPov} dataset. Main changes are related to the panel specific attributes.', fig.width = 7, fig.height = 9}
lmplot(
  stat.data = edPov,
  map.data = state_polys_table,
  map.link = c("StateAb", "ID"),
  panel.types = c("map", "dot_legend", "labels", "dot", "dot"),
  panel.data = list(NA, NA, "state", "ed", "pov"),
  ord.by = "ed",
  rev.ord = TRUE,
  grouping = 5,
  median.row = TRUE,
  colors = brewer.pal(n = 5, name = "BrBG"),
  panel.att = list(
    list(
      1,
      header = "Light Gray Means\nPreviously Displayed",
      map.all = TRUE,
      fill.regions = "aggregate",
      active.border.color = "black",
      active.border.size = 1.2,
      inactive.border.color = gray(0.7),
      inactive.border.size = 1,
      panel.width = 0.8
    ),
    list(
      2,
      point.type = 20,
      point.border = TRUE,
      point.size = 2,
      panel.width = 0.9
    ),
    list(
      3,
      header = "States",
      align = "left",
      text.size = 0.9,
      panel.width = 0.7
    ),
    list(
      4,
      header = "Percent Adults With\n4+ Years of College",
      graph.bgcolor = "lightgray",
      point.size = 1.5,
      xaxis.ticks = list(10, 20, 30, 40),
      xaxis.labels = list(10, 20, 30, 40),
      xaxis.title = "Percent"
    ),
    list(
      5,
      header = "Percent Living Below\nPoverty Level",
      graph.bgcolor = "lightgray",
      point.size = 1.5,
      xaxis.ticks = list(5, 10, 15, 20),
      xaxis.labels = list(5, 10, 15, 20),
      xaxis.title = "Percent"
    )
  )
)
```


In the final revised linked micromap plot\index{Linked micromap plot} shown in
Figure \@ref(fig:Ch2-refining6), we make three more types of changes.
First, we make use of the `labeling::extended()` function of the **labeling**\index{R Packages!labeling} R package.
This function is provided with the minimum and maximum values of a variable and the tentative
number of tic marks and tic marks labels for that variable and it then creates a vector with
near-optimal axis labels. The argument `m` is used as a guideline for the number of axis labels,
but the actual number of near-optimal axis labels may differ slightly. The reader is encouraged to experiment with
`m = 2` to `m = 6` in the R code below.

Second, we use the setting `fill.regions = "two ended"`. This setting makes most sense when
`median.row = TRUE` (as is the case here) and the focus of the maps is to 
indicate which subregions are above or below the median value
of the variable specified in the `ord.by` argument (here `ed`).
Similar to the setting `fill.regions = "aggregate"`,
the subregions from all previous perceptual groups are filled in the
subsequent perceptual groups. This filing proceeds from
the top perceptual group to the median row 
and from the bottom perceptual group to the median row
by sequentially
filling the subregions that have already been displayed on the more extreme ends.
The text in the `header` has been updated to communicate this information to the reader.

Third, we reduce the margin space between some of the columns via the
`right.margin` and `left.margin` settings. Negative values are allowed.
Fine-tuning the margin spacing and the widths of the columns can require
a few iterations. The reader always should check carefully that no identifiers
are truncated or overprinted, in particular that no letters
from the longest identifier (here `Washington D.C.`) are cut off.


```{r Ch2-refining6, fig.cap = 'Sixth (and final) refined linked micromap plot\\index{Linked micromap plot}, based on the _edPov_\\index{Datasets!edPov} dataset. Main changes are related to labeling, the coloring of perceptual groups above and below the median row, and the column spacing.', fig.width = 7, fig.height = 9}
library(labeling)

lmplot(
  stat.data = edPov,
  map.data = state_polys_table,
  map.link = c("StateAb", "ID"),
  panel.types = c("map", "dot_legend", "labels", "dot", "dot"),
  panel.data = list(NA, NA, "state", "ed", "pov"),
  ord.by = "ed",
  rev.ord = TRUE,
  grouping = 5,
  median.row = TRUE,
  colors = brewer.pal(n = 5, name = "BrBG"),
  panel.att = list(
    list(
      1,
      header = "Two-ended\nCumulative Maps",
      map.all = TRUE,
      fill.regions = "two ended",
      active.border.color = "black",
      active.border.size = 1.2,
      inactive.border.color = gray(0.7),
      inactive.border.size = 1,
      panel.width = 0.8
    ),
    list(
      2,
      point.type = 20,
      point.border = TRUE,
      point.size = 2,
      panel.width = 0.9
    ),
    list(
      3,
      header = "States",
      align = "left",
      right.margin = 0,
      left.margin = -1,
      text.size = 0.9,
      panel.width = 0.6
    ),
    list(
      4,
      header = "Percent Adults With\n4+ Years of College",
      graph.bgcolor = "lightgray",
      right.margin = 0,
      left.margin = -0.6,
      point.size = 1.5,
      xaxis.ticks = as.list(extended(
        dmin = min(edPov$ed),
        dmax = max(edPov$ed),
        m = 5
      )),
      xaxis.labels = as.list(extended(
        dmin = min(edPov$ed),
        dmax = max(edPov$ed),
        m = 5
      )),
      xaxis.title = "Percent"
    ),
    list(
      5,
      header = "Percent Living Below\nPoverty Level",
      graph.bgcolor = "lightgray",
      right.margin = 0.25,
      left.margin = -0.6,
      point.size = 1.5,
      xaxis.ticks = as.list(extended(
        dmin = min(edPov$pov),
        dmax = max(edPov$pov),
        m = 4
      )),
      xaxis.labels = as.list(extended(
        dmin = min(edPov$pov),
        dmax = max(edPov$pov),
        m = 4
      )),
      xaxis.title = "Percent"
    )
  )
)
```


What remains to be done is a summary and interpretation of the 
final revised linked micromap plot\index{Linked micromap plot} shown in Figure \@ref(fig:Ch2-refining6).
This plot shows dot plots of two statistical variables, the percentage of adults with four or more years of college
(in the fourth column) and the percentage living below poverty level (in the fifth column) 
in the 50 US states and Washington, D.C.

The percentage of adults with four or more years of college is used as the sorting variable
for the rows in the plot -- with highest percentages shown at the top and lowest percentages
shown at the bottom.
The map panel in the first column shows some noticeable, but not very strong spatial patterns.
Highest percentages of adults with four or more years of college can be found in the
northeastern states. In fact, eight of the top-10 states are located in the northeast,
with Washington, D.C., having the highest percentage with almost 40%. When looking at the other
maps above the median state (here, Arizona), additional eastern states, but also several western
states can be seen.
When looking at the two maps at the bottom of the plot, mostly southern states can be seen.
West Virginia is the state with the lowest percentage of only about 15%
of adults with four or more years of college. Overall,
primarily southern and central states can be seen in the maps below the median state.

The percentage living below poverty level (in the fifth column) shows a different pattern.
States with a high percentage of adults with four or more years of college have
a low percentage living below poverty level. 
Visually, the dots in the fourth and fifth column diverge, forming some crude reflected V-shape.
Washington, D.C., is a major outlier as it has the highest 
percentage of adults with four or more years of college (almost 40%)
but also the highest percentage living below poverty level (more than 20%).
Another interesting state is New Mexico with 
an above median percentage of adults with four or more years of college,
but with the fourth highest percentage living below poverty level (about 18%).
As expected, the overall correlation between these two variables is negative,
but given these two major outliers and several minor outliers
(such as Indiana and Nevada that both have a relatively low percentage living below poverty level
despite being among the bottom-10 states with respect to the percentage of
adults with four or more years of college), the correlation coefficient $r$ is relatively weak
only about `r round(cor(edPov$ed, edPov$pov), digits = 2)`.


## Example 2: A Linked Micromap Plot for Watersheds in West Virginia {#Ch2-Example2}


In this second linked micromap plot\index{Linked micromap plot}
example created with the **micromap**\index{R Packages!micromap} R package, we work with the
_WV_Watershed_\index{Datasets!WV\_Watershed} dataset
from the **MPIRBook**\index{R Packages!MPIRBook} R package.
We follow the four general steps outlined in Section \@ref(Ch2-Steps) again.
However, there are several differences compared to the first example in Section \@ref(Ch2-Example1).
First, we work with external shapefiles\index{Shapefiles}
that contain both, the boundary for watersheds in West Virginia and the statistical data
used in this linked micromap plot\index{Linked micromap plot}, rather than having
a separate dataset for the statistical data. 
These external shapefiles\index{Shapefiles} have to be read in as a
SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame}
that has to be split into the geographic information and the statistical data comonents.

Moreover, we introduce two new statistical displays
for the data columns of the linked micromap plot\index{Linked micromap plot} - 
a boxplot\index{Boxplot} and a dot plot with confidence bounds\index{Dot plot with confidence bounds}. 
Different arguments are used to fine-tune this linked micromap plot\index{Linked micromap plot}.
Moreover, we will demonstrate how to add an overall statistics or criteria line to the 
statistical columns.


### Identifying and Geoprocessing of Spatial Boundary Data {#Ch2-IdentifyingWV}


The _WV\_Watershed_\index{Datasets!WV\_Watershed} dataset
from the **MPIRBook**\index{R Packages!MPIRBook} R package
is stored in external shapefiles\index{Shapefiles} that can be
read in as a SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame}
via the `rgdal::readOGR()` function from the
**rgdal**\index{R Packages!rgdal} R package. It contains
25 aggregated watersheds and subbasins in West Virginia in the United States. 
These watersheds were introduced and discussed in more detail in @MPRG2016.
In addition to the geographic information, this dataset also contains
the statistical information for the linked micromap plots\index{Linked micromap plot} in this section.

Similar to the first example in Section \@ref(Ch2-Identifying), 
we verify that this object (once read into R) indeed is a
SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame}, then look
at some of the data in the `data` component of this object, and finally plot it,
as shown in Figure \@ref(fig:Ch2-WV).
The last step is the extraction of the statistical data from this
SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame} into a regular data frame.
This can be done by accessing the data via `@data` in the 
SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame}.

There are 41 variables in the resulting data frame.
In contrast to @MPRG2016, we will focus on the conductivity variables
in this linked micromap plot\index{Linked micromap plot}.

Variable names beginning with an uppercase letter, e.g., `Cond_med`, `Cond_LCB95`,
`Cond_UCB95`, are the population estimates, representing the median,
the lower 95% confidence bound, and the upper 95% confidence bound of a variable
(here, conductivity) in each of the 25 watersheds, respectively.
These will be used for the construction of 
dot plots with confidence bounds\index{Dot plot with confidence bounds}.

Variable names starting with a lowercase letter, e.g., 
`cond_min`, `condq1`, `cond_med1`, `cond_q3`, `cond_max`,
are the descriptive statistics, representing the minimum, first quartile, median,
third quartile, and maximum of a variable (here, conductivity) in each of the 25 watersheds, respectively.
These will be used for the construction of boxplots\index{Boxplot}.

(ref:Ch2-WV-cap) Map representation of the _WV_Watershed_\index{Datasets!WV\_Watershed} spatial boundary data set for the 25 watersheds in West Virginia that is used as the basis for linked micromap plots\index{Linked micromap plot} in this section.

```{r Ch2-WV, fig.cap = '(ref:Ch2-WV-cap)', fig.width = 5, fig.height = 4}
library(micromap)

wv_watershed <- readOGR(
  dsn = "data/WV_Watershed", 
  layer = "RandomWatershed2_stats_smooth", 
  verbose = FALSE
)

class(wv_watershed)

head(wv_watershed@data, n = 2)

plot(wv_watershed)

wv_data <- wv_watershed@data

names(wv_data)
```


### Linking Spatial Boundary Data and Statistical Data {#Ch2-LinkingWV}


In this step, the SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame} first is transformed
into a regular data frame for use in a linked micromap plot\index{Linked micromap plot}
via the `micromap::create_map_table()` function. We have to indicate a variable from the `data`
component of the SpatialPolygonsDataFrame\index{SpatialPolygonsDataFrame} object that 
can be used as an ID column. 
A matching ID column has to be identified in the statistical dataset. These two ID columns
have to be used for linking the two data sets. For the _USstates_\index{Datasets!USstates} dataset,
this is usually the `ST` variable that contains the 50 abbreviations for the US states (and for D.C.).
The resulting `state_polys_table` is a regular data frame that can be used for creating a
linked micromap plot\index{Linked micromap plot} in the next step.
Note that there are typically more rows in this data frame than there are spatial subregions.
This is because many spatial subregions consist of more than just one boundary, e.g., in the
presence of islands or if a land mass is split into more than one area, e.g., as in Michigan in the US.


```{r Ch2-linkingdataframesWV}
wv_polys_table <- create_map_table(
  tmp.map = wv_watershed,
  IDcolumn = "Random_Wat"
)
# ID variable is Random_Wat
head(wv_polys_table)

class(wv_polys_table)
dim(wv_polys_table)
names(wv_polys_table)
```


_edPov_\index{Datasets!edPov} is a data frame that contains
education and poverty level data for the 50 states (and Washington, D.C.).
This data frame has 51 rows, one for each of the 50 states (and for D.C.).
We have to identify one variable from this data frame that can be used for linking the
two data sets. Here, this is the `StateAb` variable.

The last expression in the following R code verifies that there is indeed at least one matching ID 
in the `state_polys_table` data frame for each ID in the _edPov_\index{Datasets!edPov} dataset.
Mismatching or missing IDs in the statistical data frame may cause
that data for some subregions do not appear
on any of the maps later on or that some subregions shown on the maps do not get colored in any of the maps.
While this often can be contributed to missing data in the statistical data frame, e.g., there is
only data for the 50 states but not for Washington, D.C., mismatching identifiers also can be
the underlying reason (e.g., if the statistical data frame uses `D.C.` as ID while
the spatial data frame uses `DC`). Here, everything is matching.


```{r Ch2-linkingdataframesWV2}
all(sort(wv_data$Random_Wat) == sort(unique(wv_polys_table$ID)))
```


### Creating a Draft Linked Micromap Plot {#Ch2-CreatingWV}


Now we can create a minimal draft linked micromap plot\index{Linked micromap plot} using the
`micromap::lmplot()` function, based on the previously created 
`state_polys_table` data frame and the _edPov_\index{Datasets!edPov} dataset.
In our function call, we only provide seven required arguments. None of these arguments has a default setting.
For all other arguments of this function, the default settings will be used here. 
The statistical data (_edPov_\index{Datasets!edPov}) is assigned to the `stat.data` argument
and the spatial boundary data (in the `state_polys_table` data frame) is assigned to the `map.data` argument.

The `map.link` argument is needed to tell the function how to link the statistical data
and the spatial boundary data. A vector with the names of the two linking variables 
identified in the previous step is needed for this argument.
The first variable name (`StateAb`) must come from the statistical data (here _edPov_\index{Datasets!edPov})
and the second variable name (`ID`) must come from the `state_polys_table` data frame that
contains the spatial boundary information.
Changing the order of these two variable names typically results in an error.

The `panel.types` and `panel.data` are closely related. As the name suggests, `panel.types` is a vector
that specifies the layout of the columns in the linked micromap plot\index{Linked micromap plot}.
Here, we have the `dot_legend` in the first column, `labels` in the second column,
two columns with statistical data represented as dot plots\index{Dot plot} (`dot`),
and the micromaps in the fifth and final (`map`) column.
This is matched with a list of data that is used for each of the five columns.
A list is necessary for this argument as some of the data itself can be lists as we will
see in some examples XXXX later on.
The `dot_legend` simply shows a plotting symbol in a certain color that represents that row in
the linked micromap plot\index{Linked micromap plot}. Thus, no further data is needed and `NA`
is assigned. `labels` requires some text argument, typically some identifiers of the subregions
in the maps. Here `state` from _edPov_ is used. 
The variables `pov` and `ed` from _edPov_ are used for the statistical displays in columns three and four.
It should be noted that all data specified in `panel.data` by default is taken from the
data frame specified in the `stat.data` argument.
The fifth and final column contains the micromaps. Their boundaries are obtained from
the `map.data` data frame. Thus, no further data has to be specified here. Instead, `NA` 
is assigned here.

Two more arguments have to be specified: `ord.by` specifies the sorting variable of the 
rows in the linked micromap plot\index{Linked micromap plot}. Here, `pov` 
from the `stat.data` argument is used. The sorting of the rows goes from smallest (on top) to
largest (at the bottom). Finally, `grouping` specifies  the number of rows
in each of the perceptual groups (here `5` rows). If a single integer value is provided, 
that value is used for all perceptual groups. If a vector of integer values is provided,
each perceptual group may have a different number of rows as 
shown in some examples XXXX later on.

The resulting draft linked micromap plot\index{Linked micromap plot} is shown in
Figure \@ref(fig:Ch2-creatingdraft). It is noteworthy that there are eleven perceptual groups
overall --- ten with five subregions and one with just one subregion. This is due to the fact
that there are 51 subregions overall: The 50 US states and Washington, D.C.
We abstain from interpreting this figure at this stage, but we will provide some
helpful interpretation once we have created the last refined version of this figure
in the next step.


```{r Ch2-creatingdraftWVdot, fig.cap = 'Draft linked micromap plot\\index{Linked micromap plot}, based on the _edPov_\\index{Datasets!edPov} dataset.', fig.width = 7, fig.height = 7}
# Chunk 2. code below for publication plot of conductivity descriptive statistics
# and estimates together
### add.line=129 for estimated median conductvity
# Draft plot
lmplot(
  stat.data = wv_data,
  map.data = wv_polys_table,
  map.link = c("Random_Wat", "ID"),
  panel.types = c("dot_legend", "labels", "dot", "dot", "map"),
  panel.data = list(NA, "Random_Wat", "cond_med1", "Cond_med", NA),
  ord.by = "Cond_med", 
  rev.ord = TRUE,
  grouping = 5,
  median.row = FALSE
)
```


```{r Ch2-shortennamesWV}
# shorten names of Random_Watersheds that combine 7 smaller HUC08s with larger
# neighboring HUC08s in WV Sampling Frame
wv_data$short_name <- as.factor(wv_data$Random_Wat)

levels(wv_data$short_name) <- list(
  "S. Br. Potomac" = "South Branch Potomac",
  "N. Br. Potomac" = "North Branch Potomac",
  "Cacapon/Shen." = "Cacapon/Shenandoah Hardy",
  "Potomac/Shen." = "Potomac Direct Drains/Shenandoah Jefferson",
  "U. New/James" = "Upper New/James",
  "Tygart Valley" = "Tygart Valley",
  "West Fork" = "West Fork",
  "Mononga./Dunk." = "Monongahela/Dunkard",
  "Cheat/Youg." = "Cheat/Youghiogheny",
  "U. Ohio N./S." = "Upper Ohio North/Upper Ohio South",
  "M. Ohio N." = "Middle Ohio North",
  "M. Ohio S." = "Middle Ohio South",
  "Little Kanawha" = "Little Kanawha",
  "Greenbrier" = "Greenbrier",
  "L. New" = "Lower New",
  "Gauley" = "Gauley",
  "U. Kanawha" = "Upper Kanawha",
  "Elk" = "Elk",
  "L. Kanawha" = "Lower Kanawha",
  "Coal" = "Coal",
  "U. Guyandotte" = "Upper Guyandotte",
  "L. Guyandotte" = "Lower Guyandotte",
  "Tug Fork" = "Tug Fork",
  "Big Sandy/L. Ohio" = "Big Sandy/Lower Ohio",
  "Twelvepole" = "Twelvepole"
)
```


```{r Ch2-creatingdraftWV, fig.cap = 'Draft linked micromap plot\\index{Linked micromap plot}, based on the _edPov_\\index{Datasets!edPov} dataset.', fig.width = 7, fig.height = 7}
# Chunk 2. code below for publication plot of conductivity descriptive statistics
# and estimates together
### add.line=129 for estimated median conductvity
# Draft plot
lmplot(
  stat.data = wv_data,
  map.data = wv_polys_table,
  map.link = c("Random_Wat", "ID"),
  panel.types = c("dot_legend", "labels", "box_summary", "dot_cl", "map"),
  panel.data = list(
    NA,
    "short_name",
    list("cond_min", "condq1", "cond_med1", "condq3", "cond_max"),
    list("Cond_med", "Cond_LCB95", "Cond_UCB95"),
    NA
  ),
  ord.by = "Cond_med", 
  rev.ord = TRUE,
  grouping = 5,
  median.row = FALSE
)
```


### Refining the Linked Micromap Plot {#Ch2-RefiningWV}


We continue with the draft linked micromap plot\index{Linked micromap plot} from the previous
step and refine it in multiple small steps. This refinement should only be started once
functional R code has been obtained in the previous step.

First, we get rid of the eleventh perceptual group (with just one subregion) and introduce
a median row via `median.row = TRUE` instead. A median row is often a good solution if the 
number of subregions is odd such as for the 51 states (and D.C.) here.
Here, Wyoming is the state shown in the median row. It has the 
26$^{th}$ highest (or lowest) value for the sorting variable, i.e., `pov`.
It does not appear in a map by itself, but rather is added to the perceptual
groups above and below the median row in a neutral color, 
thus increasing the number of subregions shown in each of these two maps by one (i.e., six here)
Also, we reverse the sorting order via `rev.ord = TRUE`. 
Now the sorting of the rows goes from largest (on top) to
smallest (at the bottom).
The resulting linked micromap plot\index{Linked micromap plot} is shown in
Figure \@ref(fig:Ch2-refining1).
While we keep `pov` as the sorting variable in this first refined version,
the reader is encouraged to use `ed` as sorting variables and see how
the spatial patterns highlighted in the maps change. This can be done
for the original sorting order or for the reversed sorting order.
The rows can even be sorted by `region` or alphabetically 
by `state` or `StateAb` even though such an alphabetical sorting
in most cases is not very meaningful, at least not for this specific example.


```{r Ch2-refining1WV, fig.cap = 'First refined linked micromap plot\\index{Linked micromap plot}, based on the _edPov_\\index{Datasets!edPov} dataset. Main changes are the introduction of a median row instead of the eleventh preceptual group and the reverse ordering of the `pov` data in the first statistical column.', fig.width = 7, fig.height = 7}
# Refined Plot
lmplot(
  stat.data = wv_data,
  map.data = wv_polys_table,
  map.link = c("Random_Wat", "ID"),
  panel.types = c("dot_legend", "labels", "box_summary", "dot_cl", "map"),
  panel.data = list(
    NA,
    "short_name",
    list("cond_min", "condq1", "cond_med1", "condq3", "cond_max"),
    list("Cond_med", "Cond_LCB95", "Cond_UCB95"),
    NA
  ),
  ord.by = "Cond_med", 
  rev.ord = TRUE,
  grouping = 5, 
  plot.pGrp.spacing = 1.2,
  median.row = FALSE,
  colors = brewer.pal(n = 5, name = "YlGnBu")[5:1],
  panel.att = list(
    list(
      1,
      panel.width = 1.6, 
      point.type = 15, 
      point.size = 1.2,
      point.border = TRUE
    ),
    list(
      2,
      header = "WVDEP\nSubbasins", 
      panel.width = 1.25, # increased panel.width to 1.1 from .8
      align = "left", 
      text.size = 0.8
    ),
    list(
      3,
      header = "Observed Conductivity\n4th Order Streams or Less",
      graph.bgcolor = "lightgray",
      graph.bar.size = 0.7,
      xaxis.ticks = c(0, 1000, 2000, 3000, 4000, 5000),
      xaxis.labels = c(0, 1, 2, 3, 4, 5),
      xaxis.labels.size = 1.25,
      xaxis.title = "Conductivity [in 1,000]", 
      xaxis.title.size = 1.25, 
      panel.width = 2,
      right.margin = 0,
      left.margin = -0.5
    ),
    list(
      4,
      header = "Estimated Conductivity\n4th Order Streams or Less",
      graph.bgcolor = "lightgray",
      xaxis.ticks = list(0, 150, 300, 450, 600, 750),
      xaxis.labels = list(0, 150, 300, 450, 600, 750),
      xaxis.labels.size = 1.25,
      xaxis.title = "Conductivity",
      add.line = 129, 
      add.line.col = "black", 
      add.line.typ = "dashed",
      xaxis.title.size = 1.25, 
      panel.width = 2,
      right.margin = 0.1,
      left.margin = -0.5
    ),
    list(
      5,
      header = "Light Gray Means\nHighlighted Above",
      inactive.border.color = gray(0.7), 
      inactive.border.size = 1.5,
      panel.width = 1.5
    )
  )
)
```

What remains to be done is a summary and interpretation of the 
final revised linked micromap plot\index{Linked micromap plot} shown in Figure \@ref(fig:Ch2-refining6).
This plot shows dot plots of two statistical variables, the percentage of adults with four or more years of college
(in the fourth column) and the percentage living below poverty level (in the fifth column) 
in the 50 US states and Washington, D.C.

The percentage of adults with four or more years of college is used as the sorting variable
for the rows in the plot -- with highest percentages shown at the top and lowest percentages
shown at the bottom.
The map panel in the first column shows some noticeable, but not very strong spatial patterns.
Highest percentages of adults with four or more years of college can be found in the
northeastern states. In fact, eight of the top-10 states are located in the northeast,
with Washington, D.C., having the highest percentage with almost 40%. When looking at the other
maps above the median state (here, Arizona), additional eastern states, but also several western
states can be seen.
When looking at the two maps at the bottom of the plot, mostly southern states can be seen.
West Virginia is the state with the lowest percentage of only about 15%
of adults with four or more years of college. Overall,
primarily southern and central states can be seen in the maps below the median state.

The percentage living below poverty level (in the fifth column) shows a different pattern.
States with a high percentage of adults with four or more years of college have
a low percentage living below poverty level. 
Visually, the dots in the fourth and fifth column diverge, forming some crude reflected V-shape.
Washington, D.C., is a major outlier as it has the highest 
percentage of adults with four or more years of college (almost 40%)
but also the highest percentage living below poverty level (more than 20%).
Another interesting state is New Mexico with 
an above median percentage of adults with four or more years of college,
but with the fourth highest percentage living below poverty level (about 18%).
As expected, the overall correlation between these two variables is negative,
but given these two major outliers and several minor outliers
(such as Indiana and Nevada that both have a relatively low percentage living below poverty level
despite being among the bottom-10 states with respect to the percentage of
adults with four or more years of college), the correlation coefficient $r$ is relatively weak
only about `r round(cor(edPov$ed, edPov$pov), digits = 2)`.


## Further Reading {#Ch2-FurtherReading}


Suggestions: 
For a journal, use plot.height=9, plot.width=6.8.
For a poster, use plot.height=6, plot.width=9.



Introduce cross-references to other chapters, e.g., Chapter \@ref(Ch1) and Chapter \@ref(Ch3),
where related work and further examples can be found in this book that match the content of this
chapter, that follow up on this chapter, or that are a prerequisite of this chapter.

Also, do some scientific literature review here that is specific to your chapter.
Where has this R package been introduced and used before, where have other plot types
or different countries been used in micromaps, what were other applications 
of micromaps that are related to the title and content of your chapter, etc.?


\printbibliography[segment=\therefsegment,heading=subbibliography]
