# Linked Micromap Plots via the **micromapST** R Package {#Ch3}


\chapterauthor{Linda Williams Pickle, James Blackwood Pearson, Jr., Daniel B. Carr}


Similar to Chapter \@ref(Ch2), the **micromapST**\index{R Packages!micromapST} R package [@CP2015CRAN],
accessible at https://cran.r-project.org/web/packages/micromapST/index.html,
will be introduced in this chapter. The reader will learn how to create a 
basic linked micromap plot\index{Linked micromap plot} via this R package.  
Although the default parameter settings give publication-ready output,
details will be provided on additional display options for U.S. states and other geospatial
boundary files.


## Introduction {#Ch3-Introduction}


Linked micromap plot \index{Linked micromap plot} designs can serve many purposes, including communicating to audiences in different application domains and identifying patterns useful in exploratory data analysis. Users ranging from the general public to sophisticated data analysts can understand these plots. The design removes the technical detail of plot creation from the viewer, allowing the data patterns to stand out visually. 

The **micromapST**\index{R Packages!micromapST} package simplifies creation of the micromap plot by incorporating default plot parameters that produce publication-ready output for most applications. Furthermore, its execution is quick because it plots the micromaps using fundamental R functions rather than calling another plotting package.

As noted in Chapter \@ref(Ch1), the linked micromap plot \index{Linked micromap plot} design arose from Dan Carr's wish to make geospatial data available from many federal agencies more accessible to the public. The goal was primarily to communicate those statistics in a compact, easily-understood form. Up to that point, most federal statistical reports consisted of tables sorted alphabetically by the names of the areas represented. In a meeting with Tony Olsen at the Environmental Protection Agency (EPA)\index{Environmental Protection Agency} in the early 1990s, Olsen commented that he liked Dan's sorted row plot\index{Row plot} graphical design, but also wanted to see the geospatial context of the data. Dan's immediate response was the initial linked micromap design. Dan collaborated on specific geospatial graphical design questions throughout the 1990s with researchers at the Environmental Protection Agency\index{Environmental Protection Agency}, the Bureau of Labor Statistics\index{Bureau of Labor Statistics}, the National Center for Health Statistics\index{National Center for Health Statistics} and the National Cancer Institute\index{National Cancer Institute}, leading to refinements of the design.

A secondary goal of the micromap design was to serve as a tool for exploratory spatial data analyses\index{Exploratory spatial data analysis}. For example, static maps of death rates due to cancer [@MMHBF1975;@PMHHF1987;@PMHHF1990] and other causes [@PMJW96]
generated hypotheses about possible causes of the observed geospatial patterns in these rates that could subsequently be tested by epidemiologic studies. Linked micromaps provide the analyst a tool to explore visual associations between the values of the outcome variable, such as lung cancer rates, and possible risk factors for that outcome, such as smoking rates, income and education levels. Sorting on the outcome variable or a putative risk factor can reveal geospatial clusters\index{Geospatial clusters} of similar values in the small maps, identifying the most promising places for subsequent study. These patterns can also trigger further thoughts on other variables to examine. Linked micromap plots \index{Linked micromap plot} of larger geospatial units, such as U.S. states, can help analysts quickly identify broad geospatial patterns in the data and the more promising risk factors for future study. Subsequent micromap plots can drill down to smaller geographic units, such as counties, within larger units of interest.  

Early implementations of the micromap design were done using S-plus\index{S-plus} software
[@CaPi1996;@COCPC1998;@COPC1998;@COPC2000;@CWC2000].
When the National Cancer Institute\index{National Cancer Institute} (NCI) became interested in including a web-based interactive version of linked micromaps on its proposed State Cancer Profiles\index{State Cancer Profiles} website, Dan Carr worked with Sue Bell at NCI to finalize the design and functionality of the new graphical tool, which was then implemented in JAVA\index{JAVA}
[@CCBPZ2002;@CBPZL2003].
Since this JAVA\index{JAVA} applet was specifically for publicly-available cancer data and could not be changed from the NCI\index{National Cancer Institute} production website version, Carr led the efforts to convert the earlier S-plus\index{S-plus} software to R code so that refinements of the basic design could continue. 

Later, Jim Pearson collaborated with Carr to create an R package, first released in 2013. The goal of this package development was to quickly display a visually appealing linked micromap plot \index{Linked micromap plot} from a list of user-provided parameters without requiring customizing code. The package structure used a table-driven description of the desired micromap. Filling in the table cells was analogous to filling in spreadsheet cells, a task so familiar to many users at that time. That is, the user only needed to know how to read in the data, then load and execute a package in R. Pearson generalized the original R code for ease of maintenance and added checks and meaningful messages for input parameter errors. Functionality of the package was expanded by adding many new glyphs and the ability to display a variable number of areas and maps. Because the initial implementation was for a fixed design to display U.S. states, and to distinguish this package from the EPA\index{Environmental Protection Agency} version (**micromap**), we named this package **micromapST**\index{R Packages!micromapST}. This was an unfortunate choice, because users think that the package still only displays U.S. states although it has since been expanded to show data for sub-state cancer registries, selected states at the county level, several other countries and now has the ability to read in user-defined border files.

In this chapter we demonstrate how to use **micromapST**\index{R Packages!micromapST} to create a basic linked micromap plot \index{Linked micromap plot} of U.S. state data. From the default output, we will show how to add or modify display options and then will illustrate how to plot other geographic boundary sets that are included in the package. In Chapter \@ref(Ch5b), we will demonstrate the various glyphs that are currently available in the package. In Chapter \@ref(Ch4b), we illustrate a new feature that enables users to supply their own boundary files and to create other geospatial displays.


## Basic Plot Layout and Function Call {#Ch3-Basics}


We start by producing the most basic linked micromap plot \index{Linked micromap plot} from **micromapST**\index{R Packages!micromapST}. This just involves specifying the dataset and variables to use, the glyphs to use for each statistical column and the order of displayed columns. All other parameters are set to their defaults. In the next sections, we will add  display options and change the perceptual grouping of areas, using one of the growing list of built-in boundary files as an example. 

The first examples examine the growing numbers of U.S. state covid-19 cases\index{Datasets!st\_covid\_case\_rates} from February 24, 2020, to April 14, 2022. The monthly cumulative case counts, both confirmed and presumptive, are from USAFacts [@USAFacts2022]. In order to compare states with widely varying populations, we divided each count by the state's 2019 population to obtain rates representing the numbers of cases per 1000 population. The same process was used to produce cumulative case rates for each U.S. county.

We wondered if the clear differences in the emergence and later surges of covid seen among the states were associated with differences in preventive behavior, such as social distancing and mask wearing. To illustrate how this question could be examined using linked micromap plots\index{Linked micromap plot}, we merged the cumulative case rate dataset with data on mask wearing collected from the Carnegie Mellon University's Delphi Group's Facebook survey [@SRBCLRRMLFKRT2021]. The survey question in 2020 was "In the last 5 days, how often did you wear a mask in public?". We downloaded the percents by state who responded that they wore a mask most or all of the time in September, 2020. To illustrate, the merged input data for Alaska (AK) and Alabama (AL) are shown below:

```{r Ch3-readSTCaseRates}
# Input data files.
# Data Set # 1
st_covid_case_rates <- read.csv("data/STCaseRatesPer1000.csv",
  row.names = 2
)
head(st_covid_case_rates, n = 2L)
```

To create the linked micromap, the user first specifies what to display in a data frame called `panelDesc`. This data frame has a spreadsheet-like structure, with each column of the data frame defining a column on the micromap display and each row of the data frame defining a particular display parameter. For this simple example `panel_desc_covid1` is constructed with 3 columns and 3 rows. The `type` column specifies the content of the micromap from left to right with maps on the left, area i.d. next and dot plots encoding state values on the right (`type = c("map", "id", "dot")`). The `col1` column of the data frame specifies that `Nov_20_2020`, the cumulative number of covid cases on November 20, 2020, is the variable to use in the dot plot. We need to include `NA` as placeholders for the first and second columns because only the third column has data input. Although not required, we added a label for the dot plots (`lab1`). Similar to the `col1` specification, `lab1` includes an empty string (`""`) for the map and i.d. display columns, indicating that these do not have extra text. 

After setting up the display format, we call the micromapST\index{R Packages!micromapST} function within the package. For this simple example, we only need to specify the names of the relevant data frames: where the data to be plotted reside and the name of the plot description (`panel_desc_covid1`). We also add a title for clarity.
\index{Datasets!st\_covid\_case\_rates}
```{r Ch3-LMMalpha, fig.cap = 'Cumulative number of covid-19 cases reported by U.S. states on Nov. 20, 2020.', fig.width = 8, fig.height = 10}
library(micromapST)

panel_desc_covid1 <- data.frame(
  type = c("map", "id", "dot"),
  col1 = c(NA, NA, "Nov_20_2020"),
  lab1 = c("", "", "Cases/1000")
)
panel_desc_covid1

# default = alphabetic state name sort
micromapST(
  statsDFrame = st_covid_case_rates,
  panelDesc = panel_desc_covid1,
  title = c("US State Covid Cumulative Case Rates per 1000 Population, Nov. 20, 2020")
)
```

This is a nice-looking plot that would work well for the task of reading the case rates for particular states. Thin black horizontal lines delineate the perceptual groups of 5 states each and white vertical lines on the light gray background help guide the reader's eye up or down to the axis labels. The default grouping for 51 areas is 10 groups of 5 areas each with the median-valued area shown separately in the middle of the plot.

The default order of the states in a column is alphabetic by U.S. state postal codes. This order is not useful for identifying geospatial patterns in the data. Sorting the column of states by their data values, such as covid case rates on November 20, 2020, provides a more useful design. Since this variable is already included in the `panel_desc_covid1` data frame, the arguments `sortVar = "Nov_20_2020"' and 'ascend = FALSE` in the **micromapST**\index{R Packages!micromapST} function call are the only additions needed to plot the states in a descending case rate order.
\index{Datasets!st\_covid\_case\_rates}
```{r Ch3-LMMsorted, fig.cap = 'Cumulative number of covid-19 cases reported by U.S. states on Nov. 20, 2020.', fig.width = 7, fig.height = 9}
micromapST(
  statsDFrame = st_covid_case_rates,
  panelDesc = panel_desc_covid1,
  sortVar = "Nov_20_2020",
  ascend = FALSE,
  title = c("US State Covid Cumulative Case Rates per 1000 Population, Nov. 20, 2020")
)
```

Now we can see that North and South Dakota and nearby states had high case rates in the fall of 2020. Others noticed this spike in cases at that time and speculated that a huge motorcycle rally in Sturgis ND during August 2020 might have been a covid superspreader event, because so many people were crowded together and most attendees were reportedly strongly against mask mandates [@FWGWVSHPPMYBNYCELD2020].

Let's see how our linked micromap plot \index{Linked micromap plot} could explore this hypothesis. Included in our dataset is the percent who reported that they usually wore a mask in public on a Facebook survey in September 2020 as described at the beginning of this section. We chose September because it is close to the date of the motorcycle rally but before the case spike had been reported. We can add this mask-wearing variable to the plot and look for a visual pattern suggestive of a correlation or a nonlinear functional relationship with the case rate column. 

First, we define the plot (`panel_desc_covid2`) to include four columns, the fourth (rightmost) one being the mask variable `PctWoreMask0920`, which will also be shown as a dot. Then we call the plotting function, again specifying that the states be sorted in descending order by case count. Note that since the plot title has gotten long, we split the title into two character strings; the second string prints on a second title line.
\index{Datasets!st\_covid\_case\_rates}
```{r Ch3-LMMmask, fig.cap = 'Cumulative number of covid-19 cases per 1000 by state on Nov. 20, 2020, and percent who usually wore a mask in Sep. 2020.', fig.width = 7, fig.height = 9}
panel_desc_covid2 <- data.frame(
  type = c("map", "id", "dot", "dot"),
  lab1 = c("", "", "Cases/1000", "% wore mask"),
  col1 = c(NA, NA, "Nov_20_2020", "PctWoreMask0920")
)

micromapST(
  statsDFrame = st_covid_case_rates,
  panelDesc = panel_desc_covid2,
  sortVar = "Nov_20_2020",
  ascend = FALSE,
  title = c(
    "US State Covid Cumulative Case Rates, Nov. 20, 2020",
    " and percent who usually wore a mask, Sep. 2020"
  )
)
```

There appears to be a negative correlation between the two variables. For example, we can see that the Dakotas had very high covid totals by November 2020 and residents there reported relatively low mask wearing in September. This micromap supports the theory that the Sturgis motorcycle rally was a covid superspreader event, i.e., exposure in North Dakota in August led to covid infections soon after in surrounding states where most attendees came from. This example examines the influence of mask wearing on covid case rates. The states could be sorted instead by the mask-wearing variable to look for the reverse relationship. An epidemiologic study would be needed to confirm any apparent relationship between the two variables, but this example illustrates how linked micromaps can be used to quickly identify case clusters and explore suspected risk factors.


## Display Options {#Ch3-DisplayOpt}


In this section we illustrate display modifications that support communication and pattern identification. Using full state names can help those uncertain about state abbreviations. Enhancing labels support interpretation. Bringing sequences of featured states into the foreground using color fill and outline color\index{Colors!micromapST options} can reveal larger geospatial patterns. Adding a vertical reference line\index{Reference line}, such as the U.S. rate, supports state comparisons against a common standard. Adding state rate standard error intervals shows which states have rates plausibly similar to the national rate and which do not.

Most design changes are made by changing values in the `panelDesc` data frame. For the next example, we modify the default display for the same data shown in Figure \@ref(fig:Ch3-LMMmask). Since there are four columns in the display, every parameter specification must have four values, one for each column, left to right. `NA` is used as a placeholder for each column where that parameter is not applicable. First, we change the simple dot glyph for the percent who wore a mask to include a standard error bar by changing `dot` to `dotse` in the definition of the plot columns (`type`). In Chapter \@ref(Ch5b) many more glyphs will be defined. Compared to the simple dot, this glyph needs the additional information of the length of the error bar, so a parameter (`col2`) is added to point to that value in the data set. Note that this extra value is only needed for the `dotse` column, the fourth column from left to right on the plot, so the first three positions in `col2` are shown as `NA`.  

Another addition to `panelDesc` here is `refVals`, which defines a vertical dashed line in any or all of the graphical columns. In this example, we request a line to be drawn at 85% on the mask-wearing column, the reported percent wearing masks in the U.S. on that date. `refTexts` provides a label (`US %`) for the reference line\index{Reference line} legend.

Finally, we request cumulative map shading\index{Cumulative map shading} by changing `map` to `mapcum` in the column definitions (`type`). This option shades in light yellow any map areas that had been already displayed above. That is, from the top down, each map shows the active perceptual group's\index{Perceptual group} colored areas as well as the previous groups' areas shaded in yellow. This extra color\index{Colors!micromapST options} fill pops the previously-displayed states into the foreground so that any growing geospatial cluster\index{Geospatial clusters} of sorted values is evident. This alleviates the problem of missing a cluster of similar-valued states that may fall into different, but adjacent, perceptual groups. The shading can be defined to accumulate from the top down (`mapcum`) or from the extremes to the middle (`maptail`). A third option (`mapmedian`) is a simple binary shading whereby areas that have values above the median are shaded in one color in the upper maps and areas in the lower maps with values below the median are shaded in another color.

In general, options for the display of the statistical data and glyphs are specified in `panelDesc` while options for the entire micromap display are specified in the function call to **micromapST**\index{R Packages!micromapST}. In this example, the geographic identifier is full state names, rather than the default state postal codes used in the previous examples (`plotNames = "full"`).
\index{Datasets!st\_covid\_case\_rates}
```{r Ch3-LMMoptions, fig.cap = 'Cumulative number of covid-19 cases by state on Nov. 20, 2020, and percent who usually wore a mask in Sep 2020, with optional error bars, cumulative map shading and full state names.', fig.width = 7, fig.height = 9}
panel_desc_covid2 <- data.frame(
  type = c("mapcum", "id", "dot", "dotse"),
  lab1 = c("", "", "Cases/1000", "% wore mask"),
  col1 = c(NA, NA, "Nov_20_2020", "PctWoreMask0920"),
  col2 = c(NA, NA, NA, "PctMaskStderr"),
  refVals = c(NA, NA, NA, 85),
  refTexts = "US %"
)

micromapST(
  statsDFrame = st_covid_case_rates,
  panelDesc = panel_desc_covid2,
  sortVar = "Nov_20_2020",
  ascend = FALSE,
  plotNames = "full",
  title = c(
    "US State Covid Cumulative Case Rates, Nov. 20, 2020,",
    " and percent who usually wore a mask, Sep. 2020"
  )
)
```


Looking down the map column with cumulative shading from the highest to lowest case rates reveals an enlarging cluster of states that starts in the North Central region of the U.S. This is consistent with the superspreader North Dakota motorcycle rally hypothesis.  By adding an error bar to the mask-wearing dot, we can see that even though North Dakota, South Dakota and Wyoming have wide error bars, mostly due to their smaller populations, they still appear to have values lower than most other states. There appears to be a slight negative correlation between the case and mask columns. Adding the U.S. value referent line (85%) shows that nearly all of the highest 15 case rates are in states with mask wearing percents below the U.S. value and that most of the other states have mask wearing percents about the same (error bar touches the U.S. line) or slightly above the U.S. value.

In the next figure, the maps are moved to the right side of the plot by reordering the specified glyphs in `type` and the dot with error bar is changed to a horizontal bar (`bar`). Note that this requires a change in the order of values in all of the parameter lines. In addition, a second line of column labels is added (`lab2`) and the colors\index{Colors!micromapST options} are changed from the default colors to predefined shades of gray. We recommend using the default colors but sometimes publishers are unable to print color. The gray scale available for optional area and glyph shading is the recommended set of shades that maximizes readability and distinctions among the areas when colors are not available [@BHH2003; @BP2002]. The user can also specify a different color set by the `colors` parameter, defining the colors by their names or hexadecimal definitions. The only palette name recognized is `grays`. See the package documentation for details. 

There are 12 default colors\index{Colors!micromapST options} used:

-	Up to 6 for the map areas and symbols on the glyphs (red, orange, green, blue/green, lavender, magenta/brown);
-	1 for the median area (black);
-	3 foreground colors for highlighted areas on maps with cumulative shading (light yellow, light red, light blue);
-	1 for areas not referenced (lightest gray);
-	1 for non-active background areas (lighter gray).

\index{Datasets!st\_covid\_case\_rates}
```{r Ch3-LMMcolorsbar, fig.cap = 'Cumulative number of covid-19 cases by state on Nov. 20, 2020, and percent who usually wore a mask in Sep 2020, in gray scale with expanded labels and bar glyph.', fig.width = 7, fig.height = 9}
panel_desc_covid2 <- data.frame(
  type = c("id", "dot", "bar", "mapcum"),
  lab1 = c("", "Cases/1000", "% wore mask", ""),
  lab2 = c("", "Nov 20, 2020", "Sep 2020", ""),
  col1 = c(NA, "Nov_20_2020", "PctWoreMask0920", NA),
  col2 = c(NA, NA, "PctMaskStderr", NA),
  refVals = c(NA, NA, 85, NA),
  refTexts = "US %"
)

micromapST(
  statsDFrame = st_covid_case_rates,
  panelDesc = panel_desc_covid2,
  sortVar = "Nov_20_2020",
  ascend = FALSE,
  plotNames = "full",
  colors = "grays",
  title = c(
    "US State Covid Cumulative Case Rates, Nov. 20, 2020",
    " and percent who usually wore a mask, Sep. 2020"
  )
)
```

All of the examples in this book direct the micromap display to print on these book pages. However, the output could be written to a file for inclusion in other documents. For example, to write to a PDF file, open an output file by `pdf(file = 'samplefilename', width = 7, height = 8)` before the function call to **micromapST**\index{R Packages!micromapST}, then close the file by `dev.off()` after the call. `pdf` could be replaced by `png` to create that type of file or `windows` to direct the plot to a Windows display (with no filename specified). The default is to write to the open graphical window. The PDF and PNG file formats retain sufficient information to allow scaling and zooming in to see detail in the plot and so are preferred over other output formats that do not. Because the characteristics of a graphical window are much different from a file, it is difficult to predict how the window will display the micromap. Writing the micromap image to a file gives a consistent format. See Chapter \@ref(Ch4b) for an example of how to read an external image file back in for display.


## Mapping beyond U.S. states {#Ch3-Geog}


Over time, boundary files have been added to micromapST\index{R Packages!micromapST} that represent geographic units other than the original U.S. states. We refer to these as border groups. At this writing, the package includes generalized border groups for counties within the U.S. states of Kansas, Maryland, New York and Utah; the NCI\index{National Cancer Institute} cancer registries [@NCISEER2022Registries]; districts in Seoul, South Korea; provinces and municipalities in China; and the countries of Africa. We will examine education and poverty levels for the 24 Maryland counties to illustrate how to use these additional boundary files. In Chapter \@ref(Ch4b), we will demonstrate how the user may read in their own border group file to create a linked micromap.

The data frame `MDPovEd`\index{Datasets!MDPovEd} consists of various measures of education and poverty by county as reported on the Census Bureau's publicly available American Community Survey for 2016-2020 [@USCensus2022Edu;@USCensus2022Pov]. Maryland is a diverse state, with mostly agricultural jobs east of the Chesapeake Bay, mostly white-collar jobs in and around Baltimore and Washington D.C. (adjacent to Montgomery and Prince George's counties in MD), and mostly rural areas in the southern and western counties. Let's see if a linked micromap plot \index{Linked micromap plot} can discern these expected patterns by plotting the percent of adults who had at least four years of college and the percent of residents living below 150% of the federal poverty level. The cost of living in Maryland tends to be high, so the 150% criterion will capture those who are struggling to make ends meet even though they technically are not living below the poverty line.

First, the data file is read in, specifying the third column as containing the row names to match the area names in the border group dataset; these are the numeric county FIPS codes. The first two counties' data are printed for illustration. Then a simple linked micromap is requested, with full county names listed as the geographic identifier and the rows sorted by college education in descending order (`sortVar = "PctBachDegree", ascend = FALSE`). This example follows the design and display specification for Figure \@ref(fig:Ch3-LMMmask). Note that because the plot is for a border group other than U.S. states, we need to specify `bordGrp = "MarylandBG"`. 
\index{Datasets!MDPovEd}
```{r Ch3-LMMMDpoved, fig.cap = 'Percent of Maryland residents with income < 150% of the federal poverty level and percent with 4+ years of college, by county.', fig.width = 7, fig.height = 9}
library(micromapST)

MDPovEd <- read.csv("data/MDPovEducACSData20162020.csv", row.names = 3)
head(MDPovEd, n = 2L)

panel_desc_MD <- data.frame(
  type = c("mapcum", "id", "dot", "dot"),
  lab1 = c("", "", "% College", "% Poverty"),
  col1 = c(NA, NA, "PctBachDegree", "PctLess150Pov")
)

# sort by % college
micromapST(
  statsDFrame = MDPovEd,
  panelDesc = panel_desc_MD,
  rowNames = "id",
  sortVar = "PctBachDegree",
  ascend = FALSE,
  bordGrp = "MarylandBG",
  plotNames = "full",
  title = c(
    "Maryland counties: % Living < 150% of Poverty Level",
    "and % with 4+ Years of College"
  )
)
# Note 2 character strings for 2 lines of the long title
```

This plot shows the expected negative correlation between college education and income – counties with more college-educated residents tend to have a lower instance of poverty. The maps show a cluster of the highest education/lowest poverty in the DC suburbs plus Talbot County on the Eastern Shore, followed by a cluster of the Baltimore suburbs plus two more Eastern Shore counties. The cumulative shading through the first three maps defines a contiguous block of 14 counties in central Maryland with higher income and lower poverty than in the other 10 counties, which are located in western Maryland, the lower Eastern Shore, Cecil County in the northeast corner and Charles County in the southwest corner of the state. 

The 24 counties on this plot were split into the default perceptual group\index{Perceptual group} size of five counties each from the bottom and top, with the remaining four counties shown in the middle group. Perhaps the geospatial patterns would be clearer by using smaller groups so that there were fewer than four counties in the middle group. Using the same design specified by `panel_desc_MD`, the revised perceptual grouping is specified by `grpPattern = c(5, 4, 3, 3, 4, 5)` from top to bottom [@SC2008]. Note that the panel height is determined by the map size and so is not affected by the number of areas in each panel.
\index{Datasets!MDPovEd}
```{r Ch3-LMMMDpercep, fig.cap = 'Percent of Maryland residents with income < 150% of the federal poverty level and percent with 4+ years of college by county, displayed in smaller perceptual groups.', fig.width = 7, fig.height = 9}
micromapST(
  statsDFrame = MDPovEd,
  panelDesc = panel_desc_MD,
  rowNames = "id",
  sortVar = "PctBachDegree",
  ascend = FALSE,
  bordGrp = "MarylandBG",
  plotNames = "full",
  grpPattern = c(5, 4, 3, 3, 4, 5),
  title = c(
    "Maryland counties: % Living < 150% of Poverty Level",
    " and % with 4+ Years of College"
  )
)
```

Since the default grouping led to a fairly clear clustering of counties within Maryland, this revised grouping is only slightly better. 


## Summary and Further Reading {#Ch3-Summary}


In this chapter the reader has learned how to create a basic linked micromap plot \index{Linked micromap plot} using the package **micromapST**\index{R Packages!micromapST}, to enhance that plot with several display options and to use the additional boundary group files included in the package. In Chapter \@ref(Ch5b) we will describe the many graphics (glyphs) built into the package, beyond the simple ones used in this chapter to illustrate the basic plot. In Chapter \@ref(Ch4b), we will show how to read in an external geographic boundary file and to process it for use as a micromap.

The precursor to **micromapST**, written in JAVA, was implemented on the National Cancer Institute's State Cancer Profiles web site to communicate cancer data to interested readers and to allow them to quickly explore the data in conjunction with possible risk factors [@CBPZL2003;@BHPW2006]. Functions from the JAVA package were rewritten in R to support creation of many examples in Carr and Pickle's 2010 book [@CP2010]. That book provides much of the background on the three types of micromaps, such as the cognitive and data visualization principles underlying their designs. 

Version 1.0 of **micromapST** was released in 2013, with the R code and resulting examples further described in a 2015 paper [@PPC2015JSS]. This version of the package was used for data visualization classes at George Mason University in Fairfax, Virginia. Student feedback led to additional options and built-in geographic boundaries beyond U.S. states (version 1.1). Later versions added the ability to import the user's own boundary file (version 2.0) and a rewrite of code to avoid dependence on several retiring R packages (version 3.0).

The advantages of building many functions and datasets into the package include ease of use by those not familiar with R coding, freeing the user to focus on the task of examining patterns in the data. The package **micromap** uses a different approach, requiring the user to know enough about R coding to take advantage of that package's greater flexibility in plot design. Because **micromapST** draws the linked micromap using the `polygon` function in base R, with no need to call another drawing package, execution is very fast. Thus linked micromaps with different variables and display options can be quickly generated to identify the best plot to use for the purpose at hand. The disadvantage of this approach is that the user has less control of display features, although we have attempted to create publication-ready output without user modification and to offer many optional display modifications through parameter choices. The ease of use combined with speed of execution makes this an excellent tool for teaching, exploring and communicating spatial patterns and generating hypotheses for further analysis. 


The package **micromap** uses a different approach, requiring the user to know enough about R coding to take advantage of that package's greater flexibility in plot design.
