# Modifying Shapefiles for Use in Micromaps {#Ch4}


\chapterauthor{Braden Probst}


In the past, modifying shapefiles from publicly available sources such as the 
Global Administrative Areas Database (GADM - https://gadm.org/) 
has been a challenge for all but the most advanced R programmers. 
Often, there exist relatively small geographic areas that have to be enlarged 
to be visible in micromaps and possibly have to be shifted 
to a different location to become better visible. 
Similarly, regions far away from the main geographic area (such as Alaska and Hawaii 
for the United States) have to be shifted to reduce the empty space on a map. 
A recently developed (but still unpublished) R package, 
LMshapemaker (https://digitalcommons.usu.edu/etd/7751/), developed by the author of 
this chapter, will be the basis for this chapter.


## Introduction {#Ch4-Introduction}


As a reminder, see Chapter \@ref(Ch1) for general style requirements
for our `Micromap Plots in R` book. In particular, please do the following:

- Introduce meaningful labels for the sections, figures, and tables in your chapter.

- Create index entries for all R packages (such as the **micromap**\index{R Packages!micromap} R package)
and for all datasets (such as the _USstates_\index{Datasets!USstates} and _edPov_\index{Datasets!edPov} datasets)
that are used in your chapter.

- Include references for R packages and publications related to your chapter,
such as for the **micromap**\index{R Packages!micromap} [@PaOl2015] and 
**micromapST**\index{R Packages!micromapST} [@CP2015CRAN] R packages
and some micromap articles, book chapters, and books [@Carr2001;@SC2008;@CP2010].

- Also create index entries for main topics such as
linked micromap plots,\index{Linked micromap plot}
conditioned choropleth maps,\index{Conditioned choropleth map}
perceptual group,\index{Perceptual group}
color blindness,\index{Color blindness},
and quantile-quantile plot.\index{Quantile-quantile plot}


## Outline {#Ch4-Outline}


- overview of boundary files / regions that are part of the micromap R package

- use of external shapefiles

- thinning only (from within R)

- approach used for Latin American countries (from within R)

- use of the LMshapemaker R package at the command level

- the Shiny interface for LMshapemaker

- shapefile modifications beyond the LMshapemaker R package


## Main {#Ch4-Main}


Here goes the main content of your chapter. Introduce additional sections as needed.

For convenience, Figure \@ref(fig:Ch4-micromap1) shows one linked micromap plot\index{Linked micromap plot}
(which is the same as in Figure \@ref(fig:Ch1-micromap1)), but now formatted in a slightly more meaningful way.


```{r Ch4-micromap1, fig.cap = 'Here is a first micromap example for this chapter. Note that the figure is formatted in a slightly more meaningful way this time.', fig.width = 7, fig.height = 9}
library(micromap)

# initial example

data(USstates)
statePolys <- create_map_table(USstates, "ST")
data(edPov)

# basic figure 1
mmplot(
  stat.data = edPov,
  map.data = statePolys,
  panel.types = c("labels", "dot", "dot", "map"),
  panel.data = list("state", "pov", "ed", NA),
  ord.by = "pov",   
  grouping = 5, 
  median.row = TRUE,
  plot.width = 2, 
  plot.height = 6,
  map.link = c("StateAb", "ID")
)
```


## Code Chunks China {#Ch4-CodeChunksChina}


Example moved from Chapter \@ref(Ch2).


```{r Ch4-micromapChina1, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
# VI. Use of external shapefiles

library(rgeos)
library(rgdal)
library(maptools)
library(micromap)

### Read in the shapefile and simplify the polygons (run this part of the code only once!)

ChinaShapefile <- readOGR(
  dsn = "data/China_Shapefiles",
  layer = "export",
  verbose = TRUE
)

plot(ChinaShapefile)
summary(ChinaShapefile)
```


```{r Ch4-micromapChina2, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
gIsValid(spgeom = ChinaShapefile, byid = TRUE, reason = TRUE)

ChinaShapefileThin <- thinnedSpatialPoly(
  SP = ChinaShapefile,
  tolerance = 300,
  # try 100, 300, 3000, 30000
  minarea = 10000000000,
  topologyPreserve = TRUE,
  avoidGEOS = TRUE
)

ChinaShapefileThin <- gBuffer(
  spgeom = ChinaShapefileThin,
  width = 0,
  byid = TRUE
)

gIsValid(spgeom = ChinaShapefileThin, byid = TRUE, reason = TRUE)
plot(ChinaShapefileThin)

### Change tolerance to 100, 3000, and 30000 and run the the entire code again, starting with readOGR.
### Which of these is the best choice for tolerance? You may have to enlarge the map.

### If you plan to use LM plots for a research project, you may also want to
### enlarge small regions or shift far-away regions closer to the main region.
### Talk to me for further information.
```


```{r Ch4-micromapChina3, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### Read in the data

religion <- read.csv(
  file = "data/China_ReligionData.csv",
  header = TRUE
)

ChinaPolys <- create_map_table(
  tmp.map = ChinaShapefileThin,
  IDcolumn = "ename"
)

### Basic micromap plot

ChinaPlot <- mmplot(
  stat.data = religion,
  map.data = ChinaPolys,
  panel.types = c("labels", "dot", "map"),
  panel.data = list("Province", "Christianity", NA),
  ord.by = "Christianity",
  grouping = 5,
  median.row = TRUE,
  map.link = c("Province", "ID")
)
```


```{r Ch4-micromapChina4, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### Advanced micromap plot

ChinaReligion <- mmplot(
  stat.data = religion,
  map.data = ChinaPolys,
  panel.types = c(
    "map", "dot_legend", "labels",
    "dot", "dot", "dot", "dot"
  ),
  panel.data = list(
    NA,
    NA,
    "Province",
    "Christianity",
    "Buddhism",
    "Daoism",
    "Islam"
  ),
  map.link = c("Province", "ID"),
  ord.by = "Christianity",
  rev.ord = TRUE,
  grouping = 5,
  median.row = TRUE,
  plot.height = 9,
  plot.width = 9,
  colors = c("red", "orange", "green", "blue", "purple"),
  two.ended.maps = TRUE,
  map.all = TRUE,
  map.color2 = "lightgray",
  plot.header = "Religion in China",
  plot.header.size = 2,
  plot.header.color = "black",
  plot.panel.spacing = 0,
  panel.att = list(
    list(
      1,
      header = "Two-ended\nCumulative Maps",
      inactive.border.color = gray(.7),
      inactive.border.size = 1,
      panel.width = 1.2
    ),
    list(2,
      point.type = 20,
      point.size = 1.4
    ),
    list(
      3,
      header = "Provinces",
      panel.width = .9,
      align = "left",
      text.size = .8
    ),
    list(
      4,
      header = "Christianity",
      header.color = "red",
      graph.bgcolor = "lightgray",
      point.size = 1,
      xaxis.ticks = seq(0, 3000, by = 1000),
      xaxis.labels = seq(0, 3, by = 1),
      xaxis.title = "Number (Thousand)"
    ),
    list(
      5,
      header = "Buddhism",
      graph.bgcolor = "lightgray",
      point.size = 1,
      xaxis.ticks = seq(0, 3000, by = 1000),
      xaxis.labels = seq(0, 3, by = 1),
      xaxis.title = "Number (Thousand)"
    ),
    list(
      6,
      header = "Daoism",
      graph.bgcolor = "lightgray",
      point.size = 1,
      xaxis.ticks = seq(0, 2000, by = 1000),
      xaxis.labels = seq(0, 2, by = 1),
      xaxis.title = "Number (Thousand)"
    ),
    list(
      7,
      header = "Islam",
      graph.bgcolor = "lightgray",
      point.size = 1,
      xaxis.ticks = seq(0, 25000, by = 12500),
      xaxis.labels = seq(0, 25, by = 12.5),
      xaxis.title = "Number (Thousand)"
    )
  )
)
```


```{r Ch4-micromapChina5, fig.cap = 'Chunk 1.', fig.width = 7, fig.height = 9}
### Final advanced micromap plot with reduced margins

ChinaReligion <- mmplot(
  stat.data = religion,
  map.data = ChinaPolys,
  panel.types = c(
    "map", "dot_legend", "labels",
    "dot", "dot", "dot", "dot"
  ),
  panel.data = list(
    NA,
    NA,
    "Province",
    "Christianity",
    "Buddhism",
    "Daoism",
    "Islam"
  ),
  map.link = c("Province", "ID"),
  ord.by = "Christianity",
  rev.ord = TRUE,
  grouping = 5,
  median.row = TRUE,
  plot.height = 9,
  plot.width = 9,
  colors = c("red", "orange", "green", "blue", "purple"),
  two.ended.maps = TRUE,
  map.all = TRUE,
  map.color2 = "lightgray",
  plot.header = "Religion in China",
  plot.header.size = 2,
  plot.header.color = "black",
  plot.panel.spacing = 0,
  panel.att = list(
    list(
      1,
      header = "Two-ended\nCumulative Maps",
      inactive.border.color = gray(.7),
      inactive.border.size = 1,
      panel.width = 1.2
    ),
    list(2,
      point.type = 20,
      point.size = 1.4
    ),
    list(
      3,
      header = "Provinces",
      panel.width = .9,
      align = "left",
      text.size = .8
    ),
    list(
      4,
      header = "Christianity",
      header.color = "red",
      graph.bgcolor = "lightgray",
      point.size = 1,
      right.margin = 0,
      left.margin = -0.5,
      xaxis.ticks = seq(0, 3000, by = 1000),
      xaxis.labels = seq(0, 3, by = 1),
      xaxis.title = "Number (Thousand)"
    ),
    list(
      5,
      header = "Buddhism",
      graph.bgcolor = "lightgray",
      point.size = 1,
      right.margin = 0,
      left.margin = -0.5,
      xaxis.ticks = seq(0, 3000, by = 1000),
      xaxis.labels = seq(0, 3, by = 1),
      xaxis.title = "Number (Thousand)"
    ),
    list(
      6,
      header = "Daoism",
      graph.bgcolor = "lightgray",
      point.size = 1,
      right.margin = 0,
      left.margin = -0.5,
      xaxis.ticks = seq(0, 2000, by = 1000),
      xaxis.labels = seq(0, 2, by = 1),
      xaxis.title = "Number (Thousand)"
    ),
    list(
      7,
      header = "Islam",
      graph.bgcolor = "lightgray",
      point.size = 1,
      right.margin = 0.25,
      left.margin = -0.5,
      xaxis.ticks = seq(0, 25000, by = 12500),
      xaxis.labels = seq(0, 25, by = 12.5),
      xaxis.title = "Number (Thousand)"
    )
  )
)
```


## Summary and Further Reading {#Ch4-SummaryFurtherReading}


Introduce cross-references to other chapters, e.g., Chapter \@ref(Ch1) and Chapter \@ref(Ch2),
where related work and further examples can be found in this book that match the content of this
chapter, that follow up on this chapter, or that are a prerequisite of this chapter.

Also, do some scientific literature review here that is specific to your chapter.
Where has this R package been introduced and used before, where have other plot types
or different countries been used in micromaps, what were other applications 
of micromaps that are related to the title and content of your chapter, etc.?


\printbibliography[segment=\therefsegment,heading=subbibliography]

