# Advanced Shapefile Use in the **micromapST** R Package {#Ch4b}


\chapterauthor{Linda Williams Pickle, James Blackwood Pearson, Jr., Daniel B. Carr}


Similar to Chapter \@ref(Ch4), the **micromapST**\index{R Packages!micromapST} R package [@CP2015CRAN],
accessible at https://cran.r-project.org/web/packages/micromapST/index.html,
can be used for a large number of geographic regions that go beyond 
administrative boundaries for the United States and its 50 states.
The reader will learn how to create a 
linked micromap plot\index{Linked micromap plot} 
that makes use of these boundaries that are part of the 
**micromapST**\index{R Packages!micromapST} R package itself.


## Introduction {#Ch4b-Introduction}


As a reminder, see Chapter \@ref(Ch1) for general style requirements
for our `Micromap Plots in R` book. In particular, please do the following:

- Introduce meaningful labels for the sections, figures, and tables in your chapter.

- Create index entries for all R packages (such as the **micromap**\index{R Packages!micromap} R package)
and for all datasets (such as the _USstates_\index{Datasets!USstates} and _edPov_\index{Datasets!edPov} datasets)
that are used in your chapter.

- Include references for R packages and publications related to your chapter,
such as for the **micromap**\index{R Packages!micromap} [@PaOl2015] and 
**micromapST**\index{R Packages!micromapST} [@CP2015CRAN] R packages
and some micromap articles, book chapters, and books [@Carr2001;@SC2008;@CP2010].

- Also create index entries for main topics such as
linked micromap plots,\index{Linked micromap plot}
conditioned choropleth maps,\index{Conditioned choropleth map}
perceptual group,\index{Perceptual group}
color blindness,\index{Color blindness},
and quantile-quantile plot.\index{Quantile-quantile plot}


## Modifying the US shapefile {#Ch4b-USshapefile}


```{r Ch4b-modifiedUSshapefile, fig.cap = 'Shapefile modification of the US shapefile in 4 images.', fig.show = "hold", out.width = '50%', echo = FALSE}
knitr::include_graphics(c("img/USBorderGrp_p1.pdf", "img/USBorderGrp_p2.pdf"))
knitr::include_graphics(c("img/USBorderGrp_p3.pdf", "img/USBorderGrp_p4.pdf"))
```


## Modifying the Kentucky Shapefile {#Ch4b-Kentucky}


Here are the steps used to create the multiple maps in each PDF:

1. For KY only: the original 120 counties. Code will be shown to aggregate these into 15 Area Development Districts, used by many parts of the KY state government.

2. Next map is the file that first goes into the smoothing process.

3. Next map is after borders are smoothed and thinned by RMapShaper (we set the parameters to leave 2% of original polygon points).

4. Next map is after any user changes requested, e.g., moving and scaling (see Alaska) and rotating.

5. The final step builds the border file, rounding the (X,Y) coordinates which provides further smoothing (only slightly in these examples). This file is now ready for use in micromapST.


```{r Ch4b-modifiedKYshapefile, fig.cap = 'Shapefile modification of the KY shapefile in 5 images.', fig.show = "hold", out.width = '50%', echo = FALSE}
knitr::include_graphics(c("img/KYBorderGrp1_p1.pdf", "img/KYBorderGrp1_p2.pdf"))
knitr::include_graphics(c("img/KYBorderGrp1_p3.pdf", "img/KYBorderGrp1_p4.pdf"))
knitr::include_graphics("img/KYBorderGrp1_p5.pdf")
```



## Main {#Ch4b-Main}


Here goes the main content of your chapter. Introduce additional sections as needed.

For convenience, Figure \@ref(fig:Ch4b-micromap1) shows one linked micromap plot\index{Linked micromap plot}
(which is the same as in Figure \@ref(fig:Ch1-micromap1)), but now formatted in a slightly more meaningful way.


```{r Ch4b-micromap1, fig.cap = 'Here is a first micromap example for this chapter. Note that the figure is formatted in a slightly more meaningful way this time.', fig.width = 7, fig.height = 9}
library(micromap)

# initial example

data(USstates)
statePolys <- create_map_table(USstates, "ST")
data(edPov)

# basic figure 1
lmplot(
  stat.data = edPov,
  map.data = statePolys,
  panel.types = c("labels", "dot", "dot", "map"),
  panel.data = list("state", "pov", "ed", NA),
  ord.by = "pov",   
  grouping = 5, 
  median.row = TRUE,
  plot.width = 2, 
  plot.height = 6,
  map.link = c("StateAb", "ID")
)
```


## Further Reading {#Ch4b-FurtherReading}


Introduce cross-references to other chapters, e.g., Chapter \@ref(Ch1) and Chapter \@ref(Ch2),
where related work and further examples can be found in this book that match the content of this
chapter, that follow up on this chapter, or that are a prerequisite of this chapter.

Also, do some scientific literature review here that is specific to your chapter.
Where has this R package been introduced and used before, where have other plot types
or different countries been used in micromaps, what were other applications 
of micromaps that are related to the title and content of your chapter, etc.?


\printbibliography[segment=\therefsegment,heading=subbibliography]

