# Adding New Plot Types to Linked Micromap Plots (such as Arrow Plots, Line Charts, and Scatterplots) {#Ch5}


\chapterauthor{Sarah Schwartz, J{\"u}rgen Symanzik}


The primary graph types in the statistical panels that are supported by 
the micromap R package are dot plots, dot plots with confidence intervals, 
bar charts, and bar charts with confidence intervals. 
The reader will learn how to add new plot types to the repertoire of plot types, 
starting with arrow plots that show differences of two variables 
and scatterplots for two quantitative variables. 


## Introduction {#Ch5-Introduction}


As a reminder, see Chapter \@ref(Ch1) for general style requirements
for our `Micromap Plots in R` book. In particular, please do the following:

- Introduce meaningful labels for the sections, figures, and tables in your chapter.

- Create index entries for all R packages (such as the **micromap**\index{R Packages!micromap} R package)
and for all datasets (such as the _USstates_\index{Datasets!USstates} and _edPov_\index{Datasets!edPov} datasets)
that are used in your chapter.

- Include references for R packages and publications related to your chapter,
such as for the **micromap**\index{R Packages!micromap} [@PaOl2015] and 
**micromapST**\index{R Packages!micromapST} [@CP2015CRAN] R packages
and some micromap articles, book chapters, and books [@Carr2001;@SC2008;@CP2010].

- Also create index entries for main topics such as
linked micromap plots,\index{Linked micromap plot}
conditioned choropleth maps,\index{Conditioned choropleth map}
perceptual group,\index{Perceptual group}
color blindness,\index{Color blindness},
and quantile-quantile plot.\index{Quantile-quantile plot}


## Outline {#Ch5-Outline}


- overview of graph types that are part of the micromap R package

- general principles of adding new graph types


## Example 1: Adding Arrow Plots for Differences as a New Plot Type {#Ch4-Example1}

```{r Ch4-arrowplot, fig.cap = 'Arrow plot for differences.', fig.width = 7, fig.height = 9, eval = FALSE}
library(micromap)
library(ggplot2)

data(USstates)
data(edPov)

set.seed(123)
edPovnew <- edPov
edPovnew$edrandom <- rnorm(51, mean = 0, sd = 2)

myAtts <- sample_att()
myNumber <- 1
myAtts$colors <- c("red", "orange", "green", "blue")
myAtts[[myNumber]]$panel.data <- c("zero", "change")
myColors <- myAtts$colors
myColumns <- myAtts[[myNumber]]$panel.data
myTable <- data.frame(dummy = rep(1, 51))
myTable$data1 <- edPovnew$ed # myTable[, myColumns[1]]
myTable$data2 <- edPovnew$edrandom # myTable[, myColumns[2]]
myPanel <- ggplot(myTable) +
  geom_segment(
    aes(
      x = data1, 
      y = -pGrpOrd,
      xend = data2, 
      yend = -pGrpOrd, 
      colour = factor(color)
    ),
    arrow = arrow(length = unit(0.1, "cm"))
  ) +
  facet_grid(pGrp ~ .) +
  scale_colour_manual(values = myColors, guide = "none")

assimilatePlot(myPanel, myNumber, myAtts)
myPanelAtts <- standard_att()
myPanelAtts <- append(
  myPanelAtts,
  list(line.width = 1, tip.length = 1)
)

arrow_plot_att <- function() {
  myPanelAtts <- standard_att()
  myPanelAtts <- append(
    myPanelAtts,
    list(line.width = 1, tip.length = 1)
  )
}

arrow_plot_build <- function(myPanel, myNumber, myNewStats, myAtts) {
  myColors <- myAtts$colors
  myColumns <- myAtts[[myNumber]]$panel.data
  myLineWidth <- myAtts[[myNumber]]$line.width
  myTipLength <- myAtts[[myNumber]]$tip.length
  myTable$data1 <- myNewStats[, myColumns[1]]
  myTable$data2 <- myNewStats[, myColumns[2]]
  myPanel <- ggplot(myTable) +
    geom_segment(
      aes(
        x = data1, 
        y = -pGrpOrd,
        xend = data2, 
        yend = -pGrpOrd,
        colour = factor(color)
      ),
      arrow = arrow(length = unit(0.1 * myTipLength, "cm")),
      size = myLineWidth
    ) +
    facet_grid(pGrp ~ ., space = "free", scales = "free_y") +
    scale_colour_manual(values = myColors, guide = "none")
  myPanel <- assimilatePlot(myPanel, myNumber, myAtts)
}


```


```{r Ch4-creatingdraft, fig.cap = 'Draft linked micromap plot\\index{Linked micromap plot}, based on the _edPov_\\index{Datasets!edPov} dataset.', fig.width = 7, fig.height = 9}

library(micromap)
library(ggplot2)

data(USstates)
data(edPov)

state_polys_table <- create_map_table(
  tmp.map = USstates,
  IDcolumn = "ST"
)

mmplot(
  stat.data = edPov,
  map.data = state_polys_table,
  map.link = c("StateAb", "ID"),
  panel.types = c("dot_legend", "labels", "dot", "dot", "map"),
  panel.data = list(NA, "state", "pov", "ed", NA),
  ord.by = "pov",
  grouping = 5
)
```


## Example 2: Adding Scatterplots as a New Plot Type {#Ch4-Example2}



## Main {#Ch5-Main}


Here goes the main content of your chapter. Introduce additional sections as needed.

For convenience, Figure \@ref(fig:Ch5-micromap1) shows one linked micromap plot\index{Linked micromap plot}
(which is the same as in Figure \@ref(fig:Ch1-micromap1)), but now formatted in a slightly more meaningful way.


```{r Ch5-micromap1, fig.cap = 'Here is a first micromap example for this chapter. Note that the figure is formatted in a slightly more meaningful way this time.', fig.width = 7, fig.height = 9}
library(micromap)

# initial example

data(USstates)
statePolys <- create_map_table(USstates, "ST")
data(edPov)

# basic figure 1
lmplot(
  stat.data = edPov,
  map.data = statePolys,
  panel.types = c("labels", "dot", "dot", "map"),
  panel.data = list("state", "pov", "ed", NA),
  ord.by = "pov",   
  grouping = 5, 
  median.row = TRUE,
  plot.width = 2, 
  plot.height = 6,
  map.link = c("StateAb", "ID")
)
```


## Summary and Further Reading {#Ch5-SummaryFurtherReading}


Introduce cross-references to other chapters, e.g., Chapter \@ref(Ch1) and Chapter \@ref(Ch2),
where related work and further examples can be found in this book that match the content of this
chapter, that follow up on this chapter, or that are a prerequisite of this chapter.

Also, do some scientific literature review here that is specific to your chapter.
Where has this R package been introduced and used before, where have other plot types
or different countries been used in micromaps, what were other applications 
of micromaps that are related to the title and content of your chapter, etc.?


\printbibliography[segment=\therefsegment,heading=subbibliography]

