# Linked Micromap Plots for Point Locations {#Ch6}


\chapterauthor{Martin Holdrege, J{\"u}rgen Symanzik}


A few stand-alone attempts have been made in the past to create linked micromap plots 
for point locations such as climate stations and cities, rather than for areal locations 
(i.e., polygons). 
This requires that the point location is extended to a small (circular or quadratic) 
area that can be color-coded in a linked micromap plot. This chapter provides an 
overview of necessary steps to produce linked micromap plots for point locations 
to create areas of suitable sizes and to avoid overplotting of nearby point locations.

## Notes [to be deleted later]

To do:
-consider creating a function that will combine the circles with a background map
(this could be called `create_spatial_points_map()`)


## Introduction: The challenge of displaying point data using micromaps

One obstacle to creating micromaps for point locations is that the user 
must create polygons to represent the 
point locations. The more standard use of micromaps, displays information associated
with geographic areas for which the associated spatial data
likely already exists. In the case of point locations,
points need to displayed on the 
map as a circles (or other shape). The coordinates of a given point location become the center of the circle.
Once the polygons are created, maps and accompanying data can be visualized 
using the  **micromap**\index{R Packages!micromap} [@PaOl2015] package, 
which requires that each entity for which data
is displayed is a polygon on the map.

When creating the polygons (circles) for display on a
micromap, they
need to be appropriately sized and should not have too much overlap with each other. 
The locations of points that are very close to each other may need to be adjusted so they can be distinguished on the micromap.
Lastly, to create more useful micromaps, the polygons representing the point 
locations also need to be
integrated into a base map that shows relevant geographic boundaries. 

As with more traditional micromaps, values from multiple variables can be shown
for each location. As with other micromaps, there is a space constraint and in 
general we find it challenging to display more than roughly
50 point locations in one micromap. 

It is also important to consider whether a micromaps of point locations are actually 
the appropriate visualization for the data that you're trying to display. 
You should first consider whether the points represent a broader geographic area, which could be represented by larger polygons,
as is done in more traditional micromaps presented in the previous chapters. For example, in 
Section \@ref(Ch6-Utah) we provide an example where we plot data associated with County Seats in
the state of Utah. In that case there is one point location (city) per county, so it would
also be possible to make an effective micromap where instead of showing the point locations,
the corresponding counties were shown on the micromap, which would potentially require fewer steps to 
create. 


## Outline of steps to create micromaps of point locations


Broadly the following steps need be taken to display point locations
on micromaps. We will provide more detail on these steps in following
sections. 


1. The user needs to obtain coordinates of the point locations of interest along with 
the corresponding data to display.

1. Obtain geographic data for the underlying base map that the user wishes
to display. Strictly speaking, the micromaps of point locations could be shown without a base map
but in most cases we think that would make a less useful visualization. 

1. The locations of points that are close to each other may need to be adjusted. 
If points are very close together they may overlap on the final map,
making them hard or impossible distinguish from each other. Adjusting point locations does lead to some geographic 
distortion, but we have found that it is often helpful. 
We provide examples of both manual and more automated approaches to adjusting point locations. 

1. Polygons (circles, or other shape) need to be created around points. This step
is necessary because the **micromap** package was designed to show polygons,
not points on maps. As a work around, each point becomes a small polygon 
that is then visible in the resulting micromaps.
Multiple R packages offer ways to create buffers (polygons) around points,
and the `points2circles()` function (described below) provides a convenient wrapper to do this.

1. The base map and polygons representing point locations need to be 
combined into single object. This step may require re-projecting one or both datasets
into a common projection. The resulting object contains polygons
that will have data associated with them (i.e., the polygons showing the point locations),
as well as polygons that form the base map in the background. 

1. As described in section \@ref(#Ch2-Linking), the `mmplot()` function from the **micromap** package,
which creates the figure, requires a dataframe as input. 
Therefore, objects created in the previous step (which may, for example, be an object from  **sf** package)
needs to be flattened into dataframe. 

1. Create an initial simple micromap. To make trouble shooting easier, we recommend to at first not not add all
the visualization elements (e.g., error bars, labels, multiple variables). 

1. Iterate until the size of the circles representing point locations
and their level of overlap is acceptable.

1. As necessary, add additional visualization elements including labels, and adjust dimensions
of the figure until you're satisfied with the final product. 

We walk through the steps needed to make micromaps using three separate datasets. 
In Section \@ref(Ch6-Example1), we create a micromap showing
the populations of large cities across Asia. Section \@ref(Ch6-Utah) provides
a more complex example where point locations are manually adjusted to limit overlap, and section
\@ref(Ch6-Example3) provides an example where point locations are repelled
from each other using the `point_repel()` function. 

## Plot large cities in Asia {#Ch6-Example1}

In this example, we will visualize the locations of the 10 largest cities
in Asia, and their associated populations. 

First the necessary packages need to be loaded. 

```{r Ch6-libraries, message = FALSE, warning = FALSE}
library(micromap)
library(sf)
library(dplyr)
source("R/Ch6_functions.R") # maybe these will be integrated into a package
```

### Load point location data

Next we load the the point location data. This dataframe contains four
columns that we will use: `city`, `population`, `lat`, and `long`.

```{r Ch6-LoadCities}
cities <- readRDS("data/Ch6-data/asian_cities.RDS")
```

We can convert the population column to be in units of millions of people, 
so that the the numbers are more readable in the final
micromap. 

```{r Ch6-LoadPopulation}
cities$population <- cities$population/10^6 
```

Before moving on to create the micromap it can be helpful to create a preliminary figure
of the point locations.
Note, that when examining the locations of the cities in this dataset (Figure \@ref(fig:Ch6-AsiaScatterPlot)), none of the 
the cities are very close together, so when we draw circles around the point
locations we won't need to be worried about undo overlap in the final maps. 

```{r Ch6-AsiaScatterPlot, fig.cap='Latitude and longitude of 10 large cities in Asia. Note for this dataset none of the locations are very close to each other'}
plot(cities$long, cities$lat, xlab = "Longitude", ylab = 'Latitude')
```

### Generate polygons around points

The next step is to draw circles around the points. Here we 
achieve this using the `points2circles` function. This function can be supplied
a dataframe with columns that give the x and y coordinates, or a simple features ("sf") object
(from the **sf** package). 
By default the radius of the circles is 5% of the range of the x variable 
(i.e., 5% of the width of the map). 
The value passed to the radius argument can be changed to adjust the size of the circles,
The output of points2circles is an object of class `sf`. 

Usually metadata or other descriptions of the data will provide the coordinate
reference systems ("crs"). In this example we happen to know that the crs is `EPSG:4326` (this data is in not projected)
and we can specify that using the `crs` argument. Specifying the crs is especially 
important in section \@ref(Ch6-Example2Combine) where we combine this object
and a basemap. 

Note here we first convert the `cities` dataframe to an sf object and transform it
to a projection that is more suitable for mapping locations in Asia (being in a
'flat' projection has the added benefit of making `st_crop()` work smoothly in a later cropping step).

```{r Ch6-cities_sf}
cities_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326) %>% 
  st_transform(crs = 'EPSG:8859')
names(cities_sf)
```

The `sf` object that we created here has all the same columns  as the original dataset,
except the `long` and `lat`, columns have been removed, and there is now a `geometry` column which contains spatial data that defines the polygon for each row (city) in the dataset. `sf` objects
are essentially special kinds of dataframes, that are able to contain a wide variety
of spatial data (in the `geometry` column) and also contain additional spatial
metadata (see https://r.geocompx.org/ for more details). 


Now wer convert the points into circles.

```{r Ch6-CreateCircles-Asia}
circles_sf <- points2circles(cities_sf,
                             radius = 5)
names(circles_sf)
```


The defualt `mmplot()` function requires a dataframe as input, which can be created
using `create_map_table()` (see chapters
X, Y, Z). The 
`sf` object first needs to be converted to a `Spatial` object (from the `sp` package),
before it can be be passed to `create_map_table()`(this is accomplished with 
`as_Spatial`). 

```{r Ch6-CreateMapTableAsia}
circles_df <- create_map_table(
  tmp.map = as_Spatial(circles_sf),
  IDcolumn = "city"
)
```

### Create micromap

Now we have the data needed to create a rudimentary micromap. 
Here the micromap shows the locations of the cities, where each city
 is a circle. In the next section (\@ref(Ch6-Example2Combine)) we will tackle
how to add a base map, so that the maps are more readable. 

```{r Ch6-micromap1, fig.cap = 'Micromap showing the 10 largest cities in asia. Note that locations of cities are shown, but there are no country polygons to indicate the location of cities', fig.width = 4, fig.height = 3}
mmplot(stat.data=cities,
       map.data=circles_df,
       panel.types=c("labels", "dot", "map"),
       panel.data=list("city","population", NA),
       ord.by="population", grouping=5,
       map.link=c("city","ID"))

```

### Combine point locations and background map {#Ch6-Example2Combine}

To make a better micromap, the circles showing the city locations need to be combined
with a basemap.

Some data wrangling may be needed to acquire the necessary basemap (see Chapter X
for more details preparing maps for use in micromaps).

For this example were using a dataset of country polygons for the whole world.
This dataset can be downloaded using `rnaturalearth` package. 

```{r eval = FALSE}
wrld <- rnaturalearth::ne_download(scale = 'small', type = "countries", 
                                   returnclass = 'sf', category = 'cultural', load = TRUE)
```

However, for convenience we provide a copy of this dataset here (containing a subset of the columns).

```{r Ch6-wrld}
wrld <- readRDS("data/Ch6-data/world_countries.RDS")
```


Next we use the the `st_crop` function from the **sf** package so that only the region we're interested in will be shown.
We'll crop the map based on the extent of the `circles_sf` object we previously created.

But before we can do that we need transform the `wrld` object so that they have the 
same coordinate reference system (CRS)

```{r Ch6-CheckWrldCRS}
wrld <- st_transform(wrld, st_crs(circles_sf))
```

Since the two objects have the same CRS we don't need to do re-project, and can
crop the crop the `wrld` object. 


```{r Ch6-crop, warning = FALSE, message = FALSE}
asia <- st_crop(wrld, st_bbox(circles_sf))
```


When creating, micromaps the input includes an 'ID' column
that contains names that uniquely identify the polygons for which data will be shown (see Chapter X).
Polygons that are shown on the map but for which no data is shown (i.e., which form the map in the background)
have `NA` values in the 'ID' column. 
The 'ID' column in the `circles_sf` object  is `city`, and so here we add an empty
column of that name to `asia_sf`, beforing binding the datasets together.

```{r Ch6-ExtendAsia}
asia$city <- NA
comb_sf <- rbind(circles_sf['city'], asia['city'])
```

Note `comb_sf` object now contains two columns. 

```{r Ch6-CombColumns}
names(comb_sf)
```

Note that when selecting a specific column of a `sf` object using `object['columnName']`, an `sf` object is returned with the desired column as well as the `geometry` column. In other words, the `geometry` column is 'sticky' and comes along when selecting columns in this way, so that the associated spatial information isn't lost.

Next we need flatten the spatial object into a dataframe using the  so that it can be used
in the `mmplot()` function.


```{r Ch6-CreateMapTableCities}
comb_df <- create_map_table(
  tmp.map = as_Spatial(comb_sf),
  IDcolumn = "city"
)

head(comb_df, 3)
```

Note that in the `comb_df` data frame the `ID` column is `NA` for all polygons associated
with the base map, and is filled with the city name for all polygons (circles)
that represent the city point locations. 

### Micromap showing base map {#Ch6-Example1Micromap2}

Now we can create the micromap with the base map displayed. To properly
display the base map the `map.all` argument for the map panel needs to be set to `TRUE`
(in this case the maps are shown in the third panel). That way polygons that have no data
(i.e., polygons where the `ID` column is `NA`) also appear. 
Additionally, the `nodata.border.color` argument needs to be passed a color
(here we use "black"), to define the color of the base map lines. The width
of those lines is set with the `nodata.border.size` argument.

Other `mmplot` arguments used here are described in Chapters X and Y [insert refs].

```{r Ch6-micromap2, fig.cap = 'Micromap showing the 10 largest cities in asia', fig.width = 5, fig.height = 3}
mmplot(stat.data=cities,
       map.data=comb_df,
       panel.types=c("labels", "dot", "map"),
       panel.data=list("city","population", NA),
       ord.by="population", grouping=5,
       map.link=c("city","ID"),
       panel.att = list(
         list(2, panel.width = 0.5),
         list(2, panel.width = 0.75),
         list(3,
              map.all = TRUE,
              active.border.color = 'black', 
              active.border.size = 0.3,
              inactive.border.color = "gray", 
              inactive.border.size = 0.3, 
              panel.width = 1, 
              outer.hull =  FALSE,
              nodata.border.color = "black",
              nodata.border.size = 0.5))
)

# the mmplot.sf method doesn't work for our use here because NAs (ie for basemap) get
# plotted (delete later)
if(FALSE){
mmplot(map.data=bind_rows(circles_sf, asia['geometry']),
       panel.types=c("labels", "dot", "map"),
       panel.data=list("city","population", NA),
       ord.by="population", grouping=5,
       panel.att = list(
         list(2, panel.width = 0.5),
         list(2, panel.width = 0.75),
         list(3,
              map.all = TRUE,
              active.border.color = 'black', 
              active.border.size = 0.3,
              inactive.border.color = "gray", 
              inactive.border.size = 0.3, 
              panel.width = 1, 
              outer.hull =  FALSE,
              nodata.border.color = "black",
              nodata.border.size = 0.5))
)
}
```


## County seats in Utah {#Ch6-Utah}

For this example we will look at point location data in Utah. Specifically,
29 towns or cities in Utah that are the County seats of the 29 counties in the
state. 

### Load point locations and base map data

First we load the point locations, this is an `sf` object,
which provides the name of the county seat, as well as its elevation, population,
coordinates (in the `geometry` column), and the name of the county. 


```{r Ch6-ReadRDS}
seats1 <- readRDS("data/Ch6-data/utah_counties.RDS")
class(seats1)
names(seats1)
```

Next we want to load a spatial dataset from the **tigris** package that contains
polygons of counties
in Utah, which we can use as a basemap for the micromap.

```{r Ch6-UTCounites, cache = TRUE, message = FALSE, results='hide'}
county_polygon <- tigris::counties("Utah", cb = TRUE,
                                   year = 2020)
```

Before moving on need to check whether the base map
and point locations have the same coordinate reference system

```{r Ch6-CheckCountiesCRS}
st_crs(county_polygon) == st_crs(seats1)
```

Since they are different, one of the objects needs to be re-projected. 
Here we're re-projecting the `seats1` object, so that it has the same coordinate
reference systems as the `county_polygon` object. 

```{r Ch6-TransformCountiesCRS}
seats2 <- st_transform(seats1, crs = st_crs(county_polygon))
```

### Adjust point locations

Figure \@ref(fig:Ch6-CountyMap) shows that some county seats are quite close
to each other, especially in the north-central portion of the state. In many cases such as (such as the one shown in section \@ref(#Ch6-Example1)), no adjustment of point locations is needed because the points are naturally well spaced. 
However, in this cased we may want adjust the locations of some of these
points so that they are more easily visible in the final micromap. The user needs to
consider how they want to balance being able to see all points
in the micromap (i.e., so there is limited overlap), while not moving points
so far as to create a misleading visualization. We think it is also 
important that figure captions disclose when point locations have been adjusted. 


```{r Ch6-CountyMap, fig.cap="Locations of counties and county seats (points), in Utah."}
plot(county_polygon$geometry)
plot(seats2$geometry, add = TRUE, pch = 19, col = 'blue ')
```

There are multiple ways that point locations could be adjusted. A tedious, but effective,
method could be to manually choose new coordinates.

Another option, which we will use here, is to replace the coordinates of 
of some county seats with the centroid of their respective county. That way we can avoid
excessive over-plotting of the points on the micromap while still having
the points appear in the correct county. 

Here there are several county seats we may want to adjust the locations of. 
To do this we first create a vector of names of counties who's county seats we want to move.
Then for we calculate the centroids of those counties using the `st_centroid()` function.  

```{r Ch6-CountiesCentroids, warning = FALSE}
county_names <- c("Weber", "Morgan", "Davis", "Salt Lake", "Summit", "Wasatch",
                  "Box Elder")

centroids <- county_polygon %>% 
  filter(NAME %in% county_names) %>% 
  st_centroid()
```

Next we replace the coordinates of the county seats with the centroids of the county they're located in. 
Note here we're first arranging the columns that contain county names into alphabetical order in both the
`seats2` and `centroids` objects so that
that the row ordering matches and correct coordinates are replaced. 

```{r Ch6-CountiesReplacement}
seats_adjust1 <- arrange(seats2, county)

centroids <- arrange(centroids, NAME)

seats_adjust1$geometry[seats_adjust1$county %in% county_names] <- 
  centroids$geometry
```

Figure \@ref(fig:Ch6-CountyMap2) shows the effect of adjusting these points. 

```{r Ch6-CountyMap2, fig.cap="Original locations of county seats in blue, and adjusted locations (county centroids) in red."}
plot(county_polygon$geometry)
plot(seats2$geometry[seats2$county %in% county_names], add = TRUE, 
     pch = 19, col = 'blue')
plot(seats_adjust1$geometry[seats_adjust1$county %in% county_names], 
     add = TRUE, pch = 19, col = 'red')
```

We leave it as an exercise to the reader to adjust the locations of other
county seats to further improve the final map. 

### Prepare points and basemap for `mmplot()` 

Now that we have adjusted the point locations, we need to create circles around the points, that will be displayed on the micromap. 

```{r Ch6-PlotCircles, fig.cap="Circles drawn around thelocations of county seats. No that overlap between individual circles is fairly limited, because some locations have been adjusted."}
seats_circles <- points2circles(seats_adjust1, radius = 8)
plot(seats_circles$geometry)
```

Next the `seats_circles`  needs to be combined
with `county_polygons`. 

To be able to bind the two objects the `county_polygon` object needs to
have an empty county_seats column (as described in secion \@ref(#Ch6-Example2Combine))

```{r Ch6-ExtendCounties}
county_polygon$county_seat <- NA
seat_map <- rbind(county_polygon['county_seat'], 
                  seats_circles['county_seat'])
```

The `seat_map` object is simple features dataframe (class `sf`), which needs to be converted
to a `Spatial` object before it can be flattened into a regular data frame using the
`create_map_table` function.

```{r Ch6-CreateMapTableCounties}
seat_map_df <- create_map_table(
  tmp.map = as(seat_map, 'Spatial'),
  IDcolumn = "county_seat"
) 
```


### Create micromap

We now have the necessary data to create the micromap using the `mmplot` function.

Note that the stat.data argument in the `mmplot` function must be passed a dataframe, and the class of
`seats1` is `sf`. Therefore, we use the `st_drop_geometry` function to drop the geometry
column (i.e., which contains the spatial data) and convert the object to a regular dataframe. 

```{r Ch6-dropGeometry}
seats_df <- st_drop_geometry(seats1)
```


Note that, as described in section \@ref(#Ch6-Example1Micromap2), the `map.all`,
`nodata.border.color`, and `nodata.border.size` arguments are specified for the 
third panel so that the polygons that make up the basemap are plotted. See Chapters X and Y for more details
on other arguments used in the `mmplot` function.

```{r Ch6-MicromapSeat, fig.cap = "Micromap showing the elevations of county seats in Utah. Some locations have been adjusted to decrease overplotting", fig.width = 5, fig.height = 7}

mmplot(stat.data = seats_df,
       map.data = seat_map_df,
       panel.types = c("labels", "dot", "map"),
       panel.data = list("county_seat","elevation", NA),
       ord.by = "elevation", 
       grouping=5,
       map.link=c('county_seat', "ID"),
       panel.att = list(
         list(1, header = "County Seat", 
              panel.width = 1),
         list(2, header = "Elevation",
              xaxis.title = "feet",
              panel.width = 1.8),
         list(3, header = "Light Gray Means\nHighlighted Above",
              map.all = TRUE,
              active.border.color = 'black', 
              active.border.size = 0.3,
              inactive.border.color = "gray", 
              inactive.border.size = 0.3, 
              panel.width = 1, 
              outer.hull =  FALSE,
              nodata.border.color = "black", 
              nodata.border.size = 0.5,
              outer.hull.size = 0.5))
)
```

## Premier league football stadiums {#Ch6-Example3}

For this example we will be using using locations of premier league football 
stadiums. A challenge with this data set is being able to properly visualize multiple
locations that are very close together, such as the several stadiums located
in London. This is likely to be a common issue that a
user will run into, where locations are clustered in, for example, urban areas.

Here we show a micromap with the actual stadium locations plots, and then create a second version of the micromap where we use a more automated approach to repel locations of stadiums that are too close together. 

### Obtain point locations

The data set contains the name of the stadium, the team that plays there,
the capacity of the stadium, number of total spectators over the 2018 season,
the average number of spectators at a match, and the
coordinates (`geometry` column).

```{r Ch6-ReadStadiums}
stadiums <- readRDS("data/Ch6-data/stadiums.RDS")
names(stadiums)
```

Next we "draw" circles around the points. 

```{r Ch6-CreateCirclesStadiums}
stadium_circles <- points2circles(stadiums, radius = 8)
```

### Acquire the underlying map

We can extract a basemap of the UK, and then crop it to the extent we are 
interested (here we are only keeping the `geometry` information because we're not
interested in any of the associated country information).

```{r Ch6-ExtractUKBasemap}
uk_map <- wrld[wrld$NAME == "United Kingdom", ]
uk_map<- st_crop(uk_map['geometry'], 
                 c(xmin = -5, ymin = 50, xmax = 0, ymax = 56))

```

Now add the empty "ID" column. 

```{r Ch6-TransformUKBasemap}
uk_map$stadium <- NA
```


### Combine with points

Compare CRS

```{r Ch6-CheckCRSStadiums}
st_crs(uk_map) == st_crs(stadium_circles)
```
```{r}
uk_map <- st_transform(uk_map, st_crs(stadium_circles))
```


Next we bind the the base map and the point locations together.

```{r Ch6-CreateMapTableExtendedStadiums}
stadiums_map1 <- rbind(uk_map['stadium'], stadium_circles['stadium'])

stadiums_map_df <- create_map_table(
  tmp.map = as(stadiums_map1, 'Spatial'),
  IDcolumn = "stadium"
)
```

### Create initial micromap

Here we create a micromap showing stadium locations that are not adjusted 
(i.e. still overlapping).

```{r Ch6-micromap4, fig.cap = 'Micromap showing stadium locations, note that these maps aren not very readable due to the overlap in locations', fig.width = 5, fig.height = 7}
mmplot(stat.data=st_drop_geometry(stadiums),
       map.data=stadiums_map_df,
      panel.types=c("labels", "dot", "map"),
       panel.data=list("stadium","capacity", NA),
       ord.by="capacity", grouping=5,
       map.link=c('stadium', "ID"),
       panel.att = list(
         list(1, panel.width = 1.8),
         list(3,
              map.all = TRUE,
              active.border.color = 'black', 
              active.border.size = 0.3,
              inactive.border.color = "gray", 
              inactive.border.size = 0.3, 
              panel.width = 1, 
              outer.hull =  FALSE, # shows outside of map
              # this is required to see all map borders:
              nodata.border.color = "black", 
              nodata.border.size = 0.5,
              outer.hull.size = 0.5))
)
```

### Micromap with adjusted stadium locations

The the `point_repel()` function can be used to repel points that are too 
close together, so that in this case the cluster of stadiums in London
don't all over-plot each other. Here some of the default arguments in `point_repel()` 
were changed (`rep.fact` & `adj.max`) to increase the distance between points.
The default arguments may need to be adjustment to obtain satisfactory repulsion
between points (see function documentation for details).
Note that points that are already far apart are not moved. 

```{r Ch6-RepelPointsStadium}
stadiums_repel <- point_repel(stadiums, rep.fact = 50, adj.max = 1)
par(mfrow = c(1, 2))
plot(stadiums$geometry, main = 'Original point locations')
plot(stadiums_repel$geometry, main = 'Repelled points')
```

Now circles need to be drawn around the points.

```{r Ch6-CreateCirclesAndMapTableStadiums}
stadium_circles_repel <- points2circles(stadiums_repel, radius = 8)

stadiums_map_repel <- rbind(uk_map['stadium'], 
                            stadium_circles_repel['stadium'])

stadiums_map_repel_df <- create_map_table(
  tmp.map = as(stadiums_map_repel, 'Spatial'),
  IDcolumn = "stadium"
)
```


```{r Ch6-micromap5, fig.cap = 'Micromap showing stadium locations. Here the stadium locations have been moved slightly so that complete overlap of points does not occur. Micromaps can contain two label columns, as seen here where both the stadium name and team name are provided.', fig.width = 7.5, fig.height = 7}
mmplot(stat.data = st_drop_geometry(stadiums),
       map.data = stadiums_map_repel_df,
       panel.types = c("labels", "dot", "map", "labels"),
       panel.data = list("stadium","capacity", NA, "team"),
       ord.by = "capacity", grouping=5,
       map.link = c('stadium', "ID"),
       panel.att = list(
         list(1, header = "Stadium",
              panel.width = 1.8),
         list(2, header = "Stadium\nCapacity"),
         list(3,header = "Light Gray Means\nHighlighted Above",
              map.all = TRUE,
              active.border.color = 'black', 
              active.border.size = 0.3,
              inactive.border.color = "gray", 
              inactive.border.size = 0.3, 
              panel.width = 1, 
              outer.hull =  FALSE, 
              nodata.border.color = "black", 
              nodata.border.size = 0.5,
              outer.hull.size = 0.5),
         list(4, header =  'Team', panel.width = 1.8, align = 'left'))
)
```


## Summary and Further Reading {#Ch6-SummaryFurtherReading}


Introduce cross-references to other chapters, e.g., Chapter \@ref(Ch1) and Chapter \@ref(Ch2),
where related work and further examples can be found in this book that match the content of this
chapter, that follow up on this chapter, or that are a prerequisite of this chapter.

Also, do some scientific literature review here that is specific to your chapter.
Where has this R package been introduced and used before, where have other plot types
or different countries been used in micromaps, what were other applications 
of micromaps that are related to the title and content of your chapter, etc.?


\printbibliography[segment=\therefsegment,heading=subbibliography]

