<!-- \usepackage(hyperref) -->
<!-- \usepackage(tabulary)   # 4 lines added by LP 4/30/25 -->
<!-- \usepackage(colortbl) -->
<!-- \usepackage(lipsum) -->
<!-- \usepackage(suffix) -->
<!-- --- -->
<!-- output: -->
<!--   pdf_document: default -->
<!--   html_document: default -->
<!--   epub_document: default  # added by LP 4/30/25 -->
<!-- --- -->

# Application of Linked Micromaps in the Medical Field {#Ch12}

\chapterauthor{Linda Williams Pickle, James Blackwood Pearson, Jr.}

## Introduction {#Ch12-Introduction}

Public health studies often begin with an exploratory analysis of available data in order to generate hypotheses about a medical problem noticed in the population. Increasing disease rates over time or geospatial clusters of high rates typically trigger this sort of analysis. Questions asked in this preliminary stage include what lifestyle and environmental factors seem associated with the rates and where the problem seems to be concentrated. The latter "cluster analysis" leads epidemiologists to locations where a more in-depth study could be undertaken with sufficient numbers of cases for analysis. These questions lend themselves well to the use of linked micromaps.

In the early chapters of this book, the reader has learned about the structure and use of linked micromap plots and how to create them using two different R packages. In this chapter, the linked micromap plot will be applied to health data to illustrate how they can be useful in an exploratory analysis and subsequent communication of results to the public.

## Background {#Ch12-Background}
One of the successes in cancer screening has been the continuing decline of colorectal cancer (CRC) incidence and mortality in the U.S. since screening methods became available about 50 years ago [REF SEER]. A colonoscopy can find and remove pre-cancerous polyps before they become malignant. However, it has become apparent recently that rates among people under age 50, the recommended age to start screening, have started to rise not just in the U.S. but around the world [REF]. Researchers are puzzled by the increase. Risk factors identified for CRC occurrence in older adults include obesity, a sedentary lifestyle, smoking, excessive alcohol use and first degree relatives who also had CRC; higher death rates are seen in areas with lower screening [REF]. We will use the **micromapST**\index{R Packages!micromapST} R package [@CP2015CRAN], accessible at https://cran.r-project.org/web/packages/micromapST/index.html, to explore CRC mortality data to see if the plots can shed light on this new trend.

### Data {#Ch12-Data} 
Mortality data were obtained from the Centers for Disease Control and Prevention (CDC) online Wonder system [REF]; rates are age-adjusted using the U.S. 2000 standard population. Risk factor data were collected by CDC's Behavioral Risk Factor Surveillance System [REF] and downloaded from the National Cancer Institutes's (NCI) State Cancer Profiles system [REF]. The observed count in any strata, e.g., by state, age, sex, race or year, was suppressed in the original files if it was less than 16. Mortality data by state were available for 1999-2021, grouped into periods 1999-2001, 2002-2006, 2007-2011, 2012-2016, 2017-2021. All rates are shown as # deaths per 100,000 population. 
Risk factor data were available for 2022. Note that using this most recent risk factor data violates a major epidemiologic principle, i.e., that the putative risk factor must have occurred well prior to the initiation of the cancer. Nevertheless, these data will serve to illustrate the use of linked micromaps but we should be cautious about the interpretation of the results.

### Preliminary U.S. analysis {#Ch12-Preliminary}

First, let's look at the U.S. CRC mortality rates over time, comparing rates among those under age 50 and those 50 and older.


```{r Ch12-readMortUS}
library(micromapST)

# Input data files.
# Data Set # 1
mort_ratesUS <- read.csv("data/Ch12-data/USMortRateTrends.csv", skip = 2
)
head(mort_ratesUS, n = 2L)
```


```{r Ch12-USAgePlot, fig.cap = 'U. S. Colorectal cancer mortality rates, 1990-2022 by age group <50 and 50+', fig.width = 7, fig.height = 4}
plot(mort_ratesUS$Year,mort_ratesUS$AllMFOld,type = 'l',xlab = 'Year', ylab = 'Age-adjusted rate', ylim = c(0,85))
lines(mort_ratesUS$Year,mort_ratesUS$AllMFYoung, lty = 'dashed', col = 'blue')
legend(x = 'topright', legend = c('ages 50+','ages <50'), col = c('black','blue'), lty = c('solid','dashed'))
```

Because of the scale differences between the rates for the two age groups, we don't notice that the younger rates increased 20% while the older rates declined over 50% over this time span. From this point on, we will focus on the younger group. The numbers of deaths are relatively small, despite increasing rates, on the order of 1.5 per 100,000 population. In order to do a visual analysis of rates at the state level, how can the data be aggregated to provide reasonably stable rates and to avoid too much data suppression?

Aggregation over race/ethnic categories is a possibility, but because the non-Hispanic white category is so much larger than others, aggregate patterns will largely reflect the patterns of whites. Therefore further analyses are restricted to non-Hispanic whites.

Exploratory analyses are often conducted separately on males and females but, in the interest of avoiding small numbers in the strata, could we combine the two groups?

```{r Ch12-USSexPlot, fig.cap = 'U. S. Colorectal cancer mortality rates among white non-Hispanics ages <50, 1990-2022 for males, females and both', fig.width = 7, fig.height = 4}
plot(mort_ratesUS$Year,mort_ratesUS$WMYoung,type = 'l',lty = 'dotted', lwd = 2, col = 'blue', xlab = 'Year', ylab = 'Age-adjusted rate', ylim = c(0,3))
lines(mort_ratesUS$Year,mort_ratesUS$WFYoung, lty = 'dashed', col = 'green', lwd = 2)
lines(mort_ratesUS$Year,mort_ratesUS$WMFYoung, lty = 'solid', col = 'black', lwd = 2)
legend(x = 'bottomright', legend = c('Male','Female', 'Both'), col = c('blue','green','black'), lty = c('dotted','dashed','solid'), lwd = 2)
```

Although male rates are slightly higher than female rates, the trend lines are fairly parallel. A trend analysis showed that rates in all three groups began to increase approximately 1.6% per year since 2004, compared to a decline of 2% or more per year for whites over age 50 [REF StCaPro]. Subsequent analyses will combine white males and females.

One danger of any mortality analysis is that patterns in death rates might not reflect patterns in the underlying occurrence (incidence) due to changes in treatment or screening for that disease. For example, prostate cancer mortality rates dropped rapidly when the PSA screening test became widely available, even as incidence rates continued to rise. For CRC though, incidence rates for those under age 50 have been rising on pace with the mortality rates. This suggests that changes in screening or treatment for CRC are not the causes of the rising rates.

Researchers have not uncovered reasons for the recent increase in CRC rates among the young, although the recommended age to start screening has been lowered from 50 to 45. Risk factors for CRC occurrence in the older group include: CRC cancer in close relatives, excessive alcohol use, smoking, obesity and a sedentary lifestyle [REF NCI]. Aspirin use, hormonal therapy and pre-cancerous polyp removal seem to reduce the risk of CRC.

DATA AVAILABILITY FOR THESE RISK FACTORS
STATE DATA
SET UP LINKED MM PLOTS

```{r Ch12-readFiles}
#read csv files for mortality and risk factors, merge together.
#
#dDir <- "c:/projects/statnet/HealthChapter/"
#dDir <- "c:/MicromapBook/JuergenBase/MicromapPlotsInR-master/"
dDir <- "output/Ch12/"
File1  <- "data/Ch12-data/CRCMortData.csv"
File2  <- "data/Ch12-data/CRCRiskFactors.csv"

MortData  <- read.csv(File1, header=TRUE, stringsAsFactors = FALSE)
RiskData  <- read.csv(File2, header=TRUE, stringsAsFactors = FALSE)
MAbbr     <- MortData[,2]
RAbbr     <- RiskData[,1]
Mord      <- order(MAbbr)
Rord      <- order(RAbbr)
MortData2 <- MortData[Mord,]
RiskData2 <- RiskData[Rord,]
names(MortData2)[2] <- "MAbbr"
names(RiskData2)[1] <- "RAbbr"
row.names(MortData2) <- MortData2$MAbbr
row.names(RiskData2) <- RiskData2$RAbbr
AllData   <- cbind(MortData2,RiskData2)

stAbbr   <- MortData2$MAbbr
head(AllData)
```

COMMENT

```{r Ch12-StateLMMRates, fig.cap = 'Time Trends of Colorectal Cancer Mortality","Among Whites, Ages < 50', fig.width = 7, fig.height = 4}
# Linked MM Columns: latest time period rate, % time change from period 1 to 5, full time series
# Compute % change, 
SexBYrRates <- MortData2[,c("MAbbr","SexBYrAllAgeY","SexBYr1AgeY","SexBYr2AgeY","SexBYr3AgeY","SexBYr4AgeY","SexBYr5AgeY")]
row.names(SexBYrRates) <- stAbbr
names(SexBYrRates) <- c("Abbr","Yr0","Yr1","Yr2","Yr3","Yr4","Yr5")   # Yr0 is the ALL data over 1 to 5.
ColNames <- c("Yr0", "Yr1", "Yr2", "Yr3", "Yr4", "Yr5")

SexByrRatesArr <- array(dim=c(51,5,2),dimnames=list(stAbbr, ColNames[2:6])  )

for (ind in row.names(SexByrRatesArr)) {
     y <- as.numeric(SexBYrRates[ind,c("Yr1","Yr2","Yr3","Yr4","Yr5")])
     #names(y) <- c("Yr1","Yr2","Yr3","Yr4","Yr5")
     SexByrRatesArr[ind,,1] <- c(1,2,3,4,5)  # x value for each y
     SexByrRatesArr[ind,,2] <- y
     #print(y)
}

AllData <- cbind(MortData2,RiskData2)
AllData$Delta51  <- ( AllData$SexBYr5AgeY - AllData$SexBYr1AgeY )
AllData$PctChange <- ( AllData$SexBYr5AgeY - AllData$SexBYr1AgeY ) / AllData$SexBYr1AgeY * 100
```

```{r Ch12-setup}
# set up micromapST
#1) "Time Trends of Colorectal Cancer Mortality Among White Ages < 50"
#
#Columns:	latest time period rate, % time change from period 1 to 5, full time series.

panelHC01 <- data.frame(type=c("mapcum","id","dot","dot","ts"),
      lab1=c(NA,   NA,  "Rate",        "Pct Change",    "Time Series"),
      lab2=c(NA,   NA,  "2017-2021",   "1999-2001 to 2017-2021", "1999-2021"),
      col1=c(NA,   NA,  "SexBYr5AgeY", "PctChange",     NA),
      panelData=c(NA,NA, NA,            NA,          "SexByrRatesArr"),
      refVals=c(NA,NA,   1.8,           20,             NA),
      refTexts=c(NA,NA,  "US 2017-2021",     "US Pct Chg",   NA)
  )

#Title <- c("Time Trends of Colorectal Cancer Mortality","Among Whites, Ages < 50")

pdf(paste0(dDir,"HC01-TimeTrendsofColorectalCancerMortalityAmongWhitesAgeslessthan50SY5Rate.pdf"),height=10.5,width=7.5)

micromapST(AllData,panelHC01,sortVar="SexBYr5AgeY",ascend=FALSE, 
                   rowNames="ab", rowNamesCol="MAbbr"
  )
dev.off()

#Title <- c("Time Trends of Colorectal Cancer Mortality","Among Whites, Ages < 50")
```

RESORTED

```{r Ch12-StateLMMTrends, fig.cap = 'Time Trends of Colorectal Cancer Mortality, Sorted by percent change","Among Whites, Ages < 50', fig.width = 7, fig.height = 4}

pdf(paste0(dDir,"HC02-TimeTrendsofColorectalCancerMortalityAmongWhitesAgeslessthan50SPctChgRate.pdf"),height=10.5,width=7.5)

micromapST(AllData,panelHC01,sortVar="PctChange",ascend=FALSE, 
                   rowNames="ab", rowNamesCol="MAbbr"
 )			 

dev.off()
```

COMMENT ON THIS LMM

REDO WITH RATE & LIFESTYLE FACTORS

```{r Ch12-StateLMMRisks, fig.cap = 'Colorectal Cancer Mortality Among Whites, < 50, and related lifestyle factors', fig.width = 7, fig.height = 4}
panelHC03 <- data.frame(type=c("mapcum","id","dot",     "dot",       "dot",     "dot",     "dot"),
                        lab1=c(NA,   NA,  "Rate", "Pct Ever",  "Pct",     "Pct Binge",     "Pct"),
                        lab2=c(NA,   NA,  "2017-2021", "Smoked",    "Obese",   "Drinking",   "No Exercise"),
                        col1=c(NA,NA,"SexBYr5AgeY","PctEverSmk","PctObese","PctBinge","PctNoExer"),
                        refVals=c(NA,NA,   1.8,          NA,         NA,        NA,        NA),
                        refTexts=c(NA,NA,  "US Rate",    NA,         NA,        NA,        NA)
                   )

Title<-c("Colorectal Cancer Mortality Among Whites, < 50,","and related 
lifestyle factors.")

pdf(paste0(dDir,"HC03-ColorectalCancerMortalityAmongWhiteslessthan50andrelatedLifestyleSRate.pdf"),height=10.5,width=7.5)

micromapST(AllData,panelHC03,sortVar="SexBYr5AgeY",ascend=FALSE, 
                   rowNames="ab", rowNamesCol="MAbbr", staggerLab=TRUE    )			 

dev.off()
```

LMM FOR SCREENING AND POVERTY

```{r Ch12-StateLMMScreen, fig.cap = 'Mortality Rates and Screening for Colorectal Cancer, among Whites, ages < 50', fig.width = 7, fig.height = 4}

panelHC05 <- data.frame(type=c("mapcum","id","dot",     "dot",          "dot"),
                        lab1=c(NA,   NA,  "Rate",       "Pct CRC",          "Pct"),
                        lab2=c(NA,   NA,  "2017-2021",       "Screening",   "Poverty"),
                        col1=c(NA,   NA,  "SexBYr5AgeY", "PctCRCScreen", "PctPov"),
                        refVals=c(NA,NA,   1.8,          NA,            NA),
                        refTexts=c(NA,NA,  "US Rate",    NA,            NA)
                        
                     )

#Title <- c("Mortality Rates and Screening for Colorectal Cancer", "among Whites, ages < 50")
 
pdf(paste0(dDir,"HC05-ScreeningforColorectalCancerMortalityamongWhitesAgeslessthan50SYr5Rate.pdf"),height=10.5,width=7.5)

micromapST(AllData,panelHC05,sortVar="SexBYr5AgeY",ascend=FALSE, 
                   rowNames="ab", rowNamesCol="MAbbr", staggerLab=TRUE
  )			 

dev.off()
```
 
COMMENT ON THIS LMM

REDO WITH SORT VAR = \% CHANGE

```{r Ch12-StateLMMScreenTrends, fig.cap = 'Mortality Rates and Screening for Colorectal Cancer, among Whites, ages < 50', fig.width = 7, fig.height = 4}
#Title <- c("Screening for Colorectal Cancer Mortality","among Whites, ages < 50")
 
pdf(paste0(dDir,"HC06-ScreeningforColorectalCancerMortalityamongWhitesAgeslessthan50SChgRate.pdf"),height=10.5,width=7.5)

micromapST(AllData,panelHC05,sortVar="PctChange",ascend=FALSE, 
                   rowNames="ab", rowNamesCol="MAbbr", staggerLab=TRUE
    )			 

dev.off()
```

COMMENT
PUT IMPORTANT FACTORS ON SAME MM PLOT

```{r Ch12-StateLMMAllv1, fig.cap = 'Colorectal Cancer Mortality among Whites, ages < 50","Pct Poverty, Pct Obese, Pct No Exer', fig.width = 7, fig.height = 4}
panelHC07 <- data.frame(type=c("mapcum","id","dot",     "dot",       "dot",     "dot",       "dot"),
                        lab1=c(NA,   NA,  "Rate",       "Prd 1 to 5","Pct",     "Pct",       "Pct"),
                        lab2=c(NA,   NA,  "Yr 5",       "Pct Chg",   "Poverty", "Obesity",   "No Exercise"),
                        col1=c(NA,   NA,  "SexBYr5AgeY","PctChange", "PctPov",  "PctObese",  "PctNoExer"),
                        refVals=c(NA,NA,   1.8,         20,          NA,        NA,          NA),
                        refTexts=c(NA,NA,  "USYr5",     "USPctC",    NA,        NA,          NA)            
                        
                     )

#Title <- c("Colorectal Cancer Mortality among Whites, ages < 50","Pct Poverty, Pct Obese, Pct No Exer")
 
pdf(paste0(dDir,"HC07-ColorectalCancerMortalityamongWhitesAgeslessthan50-Pov-Obe-NoE-SYr5Rate.pdf"),height=10.5,width=7.5)

micromapST(AllData,panelHC07,sortVar="SexBYr5AgeY",ascend=FALSE, 
                   rowNames="ab", rowNamesCol="MAbbr", staggerLab=TRUE    )			 

dev.off()
```

RESORTED

```{r Ch12-StateLMMAllv2, fig.cap = 'Colorectal Cancer Mortality among Whites, ages < 50","Pct Poverty, Pct Obese, Pct No Exer', fig.width = 7, fig.height = 4}
#Title <- c("Colorectal Cancer Mortality among Whites, ages < 50","Pct Poverty, Pct Obese, Pct No Exer")
  
pdf(paste0(dDir,"HC08-ColorectalCancerMortalityamongWhitesAgeslessthan50-Pov-Obe-NoE-SYr5Rate.pdf"),height=10.5,width=7.5)
 
micromapST(AllData,panelHC07,sortVar="PctPov",ascend=FALSE, 
                    rowNames="ab", rowNamesCol="MAbbr", staggerLab=TRUE
    )			 
 
 dev.off()
```

COMMENT

```{r Ch12-StateLMMAllv3, fig.cap = 'Colorectal Cancer Mortality among Whites, ages < 50","Pct Poverty, Pct Obese, Pct No Exer', fig.width = 7, fig.height = 4}
# Title <- c("Colorectal Cancer Mortality among Whites, ages < 50","Pct Poverty, Pct Obese, Pct No Exer")
  
 pdf(paste0(dDir,"HC09-ColorectalCancerMortalityamongWhitesAgeslessthan50-Pov-Obe-NoE-SYr5Rate.pdf"),height=10.5,width=7.5)
 
 micromapST(AllData,panelHC07,sortVar="PctObese",ascend=FALSE, 
                    rowNames="ab", rowNamesCol="MAbbr", staggerLab=TRUE
    )			 
 
 dev.off()
```


## Summary and Further Reading {#Ch12-Summary}


\printbibliography[segment=\therefsegment,heading=subbibliography]
