--- 
title: "Micromap Plots in R"
subtitle: "A Step-By-Step Approach to Visualize Spatial Data via Linked Micromaps, Conditioned Micromaps, and Comparative Micromaps in R"
author: "J&uuml;rgen Symanzik"
date: "`r Sys.Date()`"
output:
  pdf_book:
    keep_tex: yes
    latex_engine: xelatex
  pdf_document: default
  html_document:
    df_print: paged
knit: "bookdown::render_book"
bibliography:
- packages.bib
- referencesMicromaps.bib
biblio-style: apa
biblatexoptions: [refsegment=chapter]
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
site: bookdown::bookdown_site
description: A book example for a Chapman & Hall book.
graphics: yes
documentclass: krantz
classoption: krantz2
header-includes:
  - \usepackage{lipsum}
  - \usepackage{suffix}
---

\renewcommand\chapterauthor[1]{\authortoc{#1}\printchapterauthor{#1}}

\makeatletter
\newcommand{\printchapterauthor}[1]{%
  {\parindent0pt\vspace*{-25pt}%
  \linespread{1.1}\large\scshape#1%
  \par\nobreak\vspace*{35pt}}}
\newcommand{\authortoc}[1]{%
  \addtocontents{toc}{\vskip-10pt}%
  \addtocontents{toc}{%
    \protect\contentsline{chapter}%
    {\hskip1.3em\mdseries\scshape\protect\scriptsize#1}{}{}}
  \addtocontents{toc}{\vskip5pt}%
}
\makeatother


```{r Setup2, include = FALSE}
# may need updated version of kableExtra from GitHub
# devtools::install_github("kupietz/kableExtra")
# see https://github.com/haozhu233/kableExtra/issues/750

options(
  htmltools.dir.version = FALSE, formatR.indent = 2, width = 80, digits = 4
)

# install the packages needed by this book; you fill out c(), e.g. c('ggplot2', 'dplyr')
lapply(c('xfun'), function(pkg) {
  if (system.file(package = pkg) == '') install.packages(pkg)
})

# hide messages and warnings from all code chunks and cache results and figures
knitr::opts_chunk$set(message = FALSE, warning = FALSE, cache = FALSE)
```


# Preface {-}


This is the preface for the `Micromap Plots in R` book.


## Why Read thisBbook {-}


You will learn a lot about all types of micromap plots.


## Structure of the Book {-}


Chapter \@ref(Ch1) provides a general overview on all types of micromap plots and provides
an overview of the following chapters of the book.

Chapter \@ref(Ch2) introduces the **micromap**\index{R Packages!micromap} R package.
Chapter \@ref(Ch3) introduces the **micromapST**\index{R Packages!micromapST} R package.


## Software Information and Conventions {-}


I used the **knitr**\index{R Packages!knitr} R package [@Xie2022knitr] and 
the **bookdown**\index{R Packages!bookdown} R package [@Xie2022bookdown] to compile the book. 
Also see the [@Xie2015] book.

My R session information is shown below:

```{r SessionInfo}
xfun::session_info()
```

Package names are in bold text (e.g., **rmarkdown**), and inline code and filenames are formatted in a typewriter font (e.g., `knitr::knit('foo.Rmd')`). Function names are followed by parentheses (e.g., `bookdown::render_book()`).


## Acknowledgments {-}


A lot of people helped me when I was writing the book. More to come later.

```{block2, type = 'flushright', html.tag = 'p'}
J&uuml;rgen Symanzik  
Logan, Utah, USA
```

